---
title: "UN_peacekeeping"
author: "Yingjie"
date: "2/23/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---



Troop and Police Contributions

  UN police and military personnel are "contributed" to serve under the blue flag from over 100 countries to carry out the mandates defined by the Security Council. Military operations aim to protect populations against threats and contribute to a secure environment. Policing is central in *preventing, mitigating and resolving violent conflict* and contributing to early peace-building and reconciliation. Uniformed personnel are contributed and deployed under several different types of arrangements.

# Package and path
```{r}
### To clear your environment 
remove(list = ls())

## tidy data
library(tidyverse)
library(lubridate)
library(viridis)
# library(xlsx)
library(readxl)
library(readr)

### set work dir
dir_this  <- dirname(rstudioapi::getSourceEditorContext()$path); dir_this


dir_root <- dirname(dir_this) %>% dirname() %>% dirname()
dir_root
source(file = paste0(dir_root, '/Code/_path of data.R'))

# setwd(dir_this)
getwd()
dir.cleaned
```


```{r Ancillary data}

### extended country-iso3 pair list for matching
# f <- paste0(dir.root, '/Data/Ancillary_Data_ISO3code_shp.RData')
# load(file = f) ## iso_eora, pop, gdp, shp, grp

f <- paste0(dir_root, '/Data/_eora_190country_iso3.xlsx')
ctr_eora <- readxl::read_xlsx(f) %>% as.data.frame()
str(ctr_eora)
```


## Functions
```{r}
not_any_na <- function(x) all(!is.na(x))
not_all_na <- function(x) any(!is.na(x))
```



# Data

## UN data

### Data on *troop-and-police-contributions*

```{r }
### data
#' download from https://peacekeeping.un.org/en/data-troop-and-police-contributions
#' The "BY GENDER CSV FILE" is very comprehensive 
#' 

yrs <- c(2015) ## used for testing
yrs <- c(2001, 2005, 2010, 2015)

getwd()
xls <- paste0(dir.peace, 'odp_contributionranks_fixCountryName.xlsx')
ctr_iso  <- readxl::read_excel(path = xls, sheet = 1) %>%
  ungroup() %>%
  distinct(ISOCode3, Contributing_Country)
names(ctr_iso) <- c('iso3', 'Contributing_Country')



# 'BY GENDER' dataset has the most comprehensive vars ------------------------------------
csv <- paste0(dir.peace, 'odp_contributionsbygender.csv')
df <- read_csv(file = csv, show_col_types = F) %>%
  mutate(date = as.Date(Last_Reporting_Date, format = '%m/%d/%Y'),
         year = year(date),
         total= Female_Personnel + Male_Personnel) %>%
  as.data.frame()

str(df)
names(df)
unique(df$Personnel_Type)


### to calculate the total contribution, by each specific date ---------------------------
df_ym <- df %>% as.data.frame() %>%
  dplyr::filter(year %in% yrs) %>%
  dplyr::select(ISOCode3, 
                Contributing_Country,
                Personnel_Type,
                Mission_Acronym,
                # Female_Personnel, Male_Personnel, 
                Last_Reporting_Date, 
                year,
                total, date) %>%
  # filter(date == as.Date('2015-12-31')) %>%
  group_by(ISOCode3, Contributing_Country, Mission_Acronym, Personnel_Type, year, date) %>%
  summarise_if(is.numeric, sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()

df_ym_agg <- df_ym %>%
  # spread(key = Personnel_Type, value = value) %>%
  # dplyr::mutate(total_contr = sum(`Experts on Mission`, Troops, `Individual Police`, `Formed Police Units`, na.rm = T)) %>%
  group_by(ISOCode3, Contributing_Country, year, date) %>%
  summarise_if(is.numeric, sum, na.rm = TRUE) %>%
  as.data.frame()
  
df_yr_byMission <- df_ym %>%
  dplyr::select(-date) %>%
  group_by(ISOCode3, Contributing_Country, Mission_Acronym, year) %>%
  summarise_if(is.numeric, sum, na.rm = TRUE) %>%
  as.data.frame()

### to list all the `Mission` names
# unique(df_yr_byMission$Mission_Acronym)


### to calculate the total contribution, by year -----------------------------------------
df_yr_agg <- df_ym %>%
  dplyr::select(-date) %>%
  group_by(ISOCode3, Contributing_Country, year) %>%
  summarise_if(is.numeric, sum, na.rm = TRUE) %>%
  ungroup()

### format for further analysis ----------------------------------------------------------
df_formated <- df_yr_agg %>%
  dplyr::select(-Contributing_Country) %>%
  # dplyr::mutate(year = ifelse(year == 2001, 2000, year)) %>%
  spread(key = year, value = total) %>%
  left_join(x = ctr_eora, y = ., by = c("iso3_eora" = "ISOCode3")) %>%
  dplyr::rename('iso3' = 'iso3_eora', 'ctr' = 'country_eora') %>%
  as.data.frame()

## Check if the column "2001" exists in the dataframe
## the earliest year is 2001. To use NA for 2000
if (!exists("2000", df_formated)) {
  # Add the column "age" with default values to the dataframe
  df_formated <- df_formated %>%
    dplyr::mutate(`2000` = NA) %>%
    dplyr::select(iso3, ctr, `2000`, everything()) %>%
    arrange(iso3)
}

f <- paste0(dir.cleaned, 'UN peacekeeping_ppl_4yrs.xlsx')
f
writexl::write_xlsx(x = df_formated, path = f)
```



  *peacekeeping operations*

  Not sure where this data comes from...
```{r eval=FALSE, include=FALSE}
### where is the data from? ------------------------------------------------------------------------
getwd()
xls <- paste0(dir.peace, 'UN peacekeeping operations.xlsx')
df  <- readxl::read_excel(path = xls, sheet = 1) 
df_op <- df %>%
  separate(col = 'MISSION NAME', into = c('name1', 'region'), 
           sep = ' in | for ', remove = F) %>%
  dplyr::select(`MISSION NAME`, name1, ACRONYM, Headquarters, Location, region, countries, everything()) %>%
  rename(from      = ACRONYM,
         from_name = name1,
         from_geo  = Headquarters, 
         to        = Location,
         to_region = region)
names(df_op)
# write.xlsx(x = dt, file = xls, sheetName = 'col_separate', row.names = F, append = T)
```



### Data from pdf 

#### data by Mission 

  The same data as `odp_contributionsbygender.csv`
  
```{r eval=FALSE, include=FALSE}
###
# if (!require("remotes")) {
#     install.packages("remotes")
# }
# # on 64-bit Windows
# remotes::install_github(c("ropensci/tabulizerjars", "ropensci/tabulizer"), INSTALL_opts = "--no-multiarch")

library(tabulizer)
pdf <- './UN peacekeeping operations/pdf_TBD/dec15_3.pdf'
tabs <- tabulizer::extract_tables(file = pdf, output = "data.frame", guess = T)

tab_df <- data.frame()

for (i in 1:length(tabs)) {
  
  tab <- tabs[i] %>% 
    as.data.frame() %>%
    select(where(not_all_na)) %>% #select(!starts_with("X"))
    # dplyr::mutate_all(na_if(.,"")) %>%
    mutate(across(where(is.character), ~na_if(., ""))) %>%
    select(where(not_all_na)) %>%
    as.data.frame()
  
  print(ncol(tab))
  
  ## We notice the number of columns in each page is different, and only 7 is correct. 
  ## - if ncol = 6, the last column should be separated into two. 
  ## - if ncol = 8, the last two columns should be combined (or added as they are numbers)
  if (ncol(tab) == 6) {
    tab <- tab %>% separate(col = ncol(.), into = c('s1', 's2'), sep = ' ')
    
    } else if (ncol(tab) == 8) {
      
      tab1 <- tab %>% 
        dplyr::mutate(
          a = .[, ncol(.)-1],
          b = .[, ncol(.)-0]) %>%
        dplyr::mutate_at(.vars = c('a', 'b'), ~gsub(",", "", x = .)) %>%
        dplyr::mutate_at(.vars = c('a', 'b'), as.numeric) %>%
        rowwise() %>%
        dplyr::mutate(s = sum(a, b, na.rm = T)) %>%
        select(-a, -b) %>%
        as.data.frame()
      n <- ncol(tab1); n
      tab1 <- tab1[, -c((n-1), (n-2))]
      tab <- tab1
      
      } else {
        tab = tab
      }
  
  print(paste0(' -- i = ', i, ' -- ', ncol(tab)))
  
  ## bind all the pages
  names(tab) <- c('Country', 'UN_Mission', 'Description', 'Male', 'Female', 'Mission', 'Total')
  tab_df <- rbind(tab_df, tab)
}


library(tidyr)

unique(tab_df$UN_Mission)  %>% sort() ## there are some non- `UN_Mission` names were mis-read into this column --> to be removed...
unique(tab_df$Description) %>% sort() ## there are some `Country` names were mis-read into this column 
unique(tab_df$Male) %>% sort()        ## there are some `Country` names were mis-read into this column
unique(tab_df$Female) %>% sort()      ## there are some `Country` names AND `UN_Mission` names were mis-read into this column

tab_df1 <- tab_df %>%
  # filter(!is.na(Country) | !is.na(Mission)) %>%
  mutate(UN_Mission2  = replace(x = UN_Mission,  grepl("Expert|Troop|Individual|Formed", UN_Mission), NA),  ## remove noise in this column
         ctr1 = replace(x = Description, grepl("Expert|Troop|Individual|Formed", Description), NA), ## to get `Country` names
         ctr1 = gsub(',|[[:digit:]]+', '', ctr1), 
         ctr2 = gsub(',|[[:digit:]]+', '', Male),
         ctrM = gsub(',|[[:digit:]]+', '', Female),
                  ) %>%
  # mutate_all(na_if(., "")) %>% 
  mutate(across(where(is.character), ~na_if(., ""))) %>% 
  dplyr::select(-Male, -Female) %>%
  as.data.frame()


unique(tab_df1$Country)
unique(tab_df1$UN_Mission2)

tab_df2 <- tab_df1 %>%
  ungroup() %>%
  ### --> fix `Country`: remove noise (i.e., any of `UN_Mission`) from `Country` names--------------
  dplyr::mutate(Country2 = ifelse(Country %in% unique(tab_df1$UN_Mission2), NA, Country),
                ctr3     = ifelse(ctrM %in% unique(tab_df1$UN_Mission2), NA, ctrM),
                m        = ifelse(is.na(ctr3), ctrM, NA),
                Country2 = ifelse(is.na(Country2), ctr1, Country2),
                Country2 = ifelse(is.na(Country2), ctr2, Country2),
                Country2 = ifelse(is.na(Country2), ctr3, Country2),
                ) %>%
  dplyr::select(Country, Country2, ctr1, ctr2, ctrM, UN_Mission, UN_Mission2, m, everything()) %>%
  dplyr::select(-Description, -ctr1, -ctr2, -ctrM, -ctr3)%>%
  ### Fill in missing values with previous or next value
  tidyr::fill(Country2) %>%
  
  ### --> fix `UN_Mission2`: get `UN_Mission` names that are in the `Country` column ---------------
  dplyr::mutate(UN_Mission2 = ifelse(is.na(UN_Mission2) & is.na(Country2), Country, UN_Mission2),
                UN_Mission2 = ifelse(is.na(UN_Mission2), m, UN_Mission2)) %>%
  dplyr::select(Country2, UN_Mission, UN_Mission2, m, everything()) %>%
  dplyr::select(-UN_Mission, -Country, -m)%>%
  as.data.frame()
  

  
tab_df3 <- tab_df2 %>%  
  
  ### --> fix `Total`: 1. some of the `Total` are in the `Mission` column -----------------------------
  mutate(Total    = ifelse(is.na(Total), Mission, Total)) %>%
  
  ### some of the `Mission` name and `Total` are in the same cell of `Total`, we should separate them
  mutate(Mission2 = gsub(',|[[:digit:]]+', '', Mission),       ## get the `Mission` name only
         m1       = gsub(',|[[:digit:]]+', '', Total),         ## get the `Mission` name only
         Total2   = as.numeric(gsub("\\D", "", Total)),
         t1       = as.numeric(gsub("\\D", "", Mission))) %>%    ## get the `Total` number only
  
  dplyr::mutate(Mission2 = dplyr::case_when(is.na(Mission2) ~ m1,
                                     Mission2 == m1  ~ Mission2,
                                     T               ~ paste0(Mission2, ' ', m1))) %>%
  as.data.frame() %>%
  ungroup() %>%
  dplyr::mutate(Mission3 = ifelse(is.na(Mission2) & !is.na(UN_Mission2), paste0('999', UN_Mission2), Mission2)) %>%
  # select(Mission, Mission2, m1, everything()) %>%
  select(-m1) %>%
  # mutate_all(na_if,"") %>% 
  mutate(across(where(is.character), ~na_if(., ""))) %>% 
  
  ### --> fix `Total`: 2. -----------------------------
  mutate(Total2 = ifelse(is.na(Total2), t1, Total2)) %>%
  
  ### some `Mission` names are in `Female` column, but should exclude values from Female if they are numbers
  # mutate(Mission2 = ifelse(is.na(Mission2), 
  #                          gsub('[[:digit:]]+', '', Female), 
  #                          Mission2)) %>%
  
  # mutate_all(na_if,"") %>% 
  # arrange(is.na(Description2)) %>%
  select(-t1, -Total, -Mission) %>% ## remove the standby data
  as.data.frame()

str(tab_df3)

### clean the data and only keep operation by county data
tab_df4 <- tab_df3 %>%
  dplyr::mutate(
    Country2 = trimws(Country2),
    Mission2 = trimws(Mission2),
    Country2 = stringr::str_squish(Country2),
    Mission2 = stringr::str_squish(Mission2)) %>%             
  dplyr::select(-UN_Mission2, -Mission3) %>%
  dplyr::filter(!is.na(Mission2) & !is.na(Total2)) %>%
  group_by(Country2, Mission2) %>%
  summarise(max = max(Total2, na.rm = T)) %>%
  ungroup() %>%
  # dplyr::mutate(check = Country2 == Mission2) %>%
  dplyr::filter(Country2 != Mission2,
                !Mission2 %in% unique(.$Country2))
  
  

str(tab_df4)
### ??? still, many countries are missing from the data ... 



### test code -------------------------------------------------------------------------------------
# tab1 <- tabs[1] %>% as.data.frame() %>% select(where(not_all_na)) %>% 
#       mutate_all(na_if,"") %>% select(where(not_all_na))
# tab2 <- tabs[2] %>% as.data.frame() %>% select(where(not_all_na)) %>% 
#       mutate_all(na_if,"") %>% select(where(not_all_na))
# tab3 <- tabs[7] %>% as.data.frame() %>% select(where(not_all_na)) %>% 
#       mutate_all(na_if,"") %>% select(where(not_all_na))
# 
# dfs <- do.call(rbind, tabs)
```




#### data by contribution 

  There is no need to use the PDF data, as we can use `odp_contributionsbygender.csv` directly. 

```{r - one example}
library(tabulizer)
pdf <- paste0(dir.peace, 'pdf/aug15_1.pdf')
tabs <- tabulizer::extract_tables(file = pdf, output = "data.frame", guess = T) 


tab_df <- data.frame()

func_list2df <- function(tabs){
  for (i in 1:length(tabs)) {
    tab <- tabs[i] %>% as.data.frame() %>% 
      select(where(not_all_na)) %>% #select(!starts_with("X"))
      # mutate_all(na_if,"") %>% 
      mutate(across(where(is.character), ~na_if(., ""))) %>% 
      select(where(not_all_na))
    
    ## *** in some cases, there is disarrangement in the 1st column! We thus need to 
    ##  drop the left NAs <in the 1st column> in a dataframe and left shifting the cleaned rows
    t1 <- t(apply(tab, 1, function(x) {
      if (is.na(x[1])) {
        y <- x[-seq_len(which.min(is.na(x))-1)]
        length(y) <- length(x)
        y
      } else x
    })) %>%
      as.data.frame() %>%
      select(where(not_all_na))
    
    
    ## *** but in some cases, there is disarrangement in the 2nd column too! We thus need to 
    ##    drop all left NAs <in the 2nd column> in a dataframe and left shifting the cleaned rows. 
    ## To do so, we first remove the 1st column as it does not have any NAs, 
    ##    then, we use the same approach as the above
    ## First, make sure there are non-NA in the 2nd column - I simply check if there are more NAs than 
    ##    the full length of the dataframe
    if (sum(is.na(t1[,2])) > nrow(t1)/2) { 
      
      t2 <- t(apply(t1[,-1], 1, function(x) {
        if (is.na(x[1])) {
          y <- x[-seq_len(which.min(is.na(x))-1)]
          length(y) <- length(x)
          y
        } else x
      })) %>%
        as.data.frame()
      
      t2 <- cbind(t1[,1], t2) ## need to add the 1st column back
        
    } else t2 <- t1
    
    
    ## clean up the table
    tab <- t2 %>% as.data.frame() %>% 
      mutate_all(as.character) %>%
      mutate_all(na_if,"") %>%
      select(where(not_all_na))
    
    # Check if the column "age" exists in the dataframe
    if (!exists("No", tab)) {
      # Add the column "age" with default values to the dataframe
      tab <- tab %>%
        dplyr::mutate(No = NA) %>%
        dplyr::select(No, everything())
    }

    
    names(tab) <- c("No", "Country", "Police", "UNMEM", "Troops", "Total" )
    
    ## bind all the pages
    tab_df <- rbind(tab_df, tab)
  }
  return(tab_df)
}


### test function 
# df <- func_list2df(tabs) %>%
#   # mutate_all(~(gsub(",", "", .))) %>% ## this also works
#   dplyr::mutate_at(.vars = 3:ncol(.), ~gsub(",", "", x = .)) %>%
#   dplyr::mutate_at(.vars = 3:ncol(.), as.numeric) %>%
#   as.data.frame()
# 
# str(df)
# 
# colSums(df[3:6], na.rm = T)
```



  There is no need to use the PDF data, as we can use `odp_contributionsbygender.csv` directly. 
```{r - loop all, eval=FALSE, include=FALSE}
library(tabulizer)
yr <- '02'
pdfs <- list.files(path = paste0(dir.peace, 'pdf/'), pattern = paste0('*', yr, '_1.pdf$'), full.names = T); pdfs

for (pdf in pdfs) {
  tabs <- tabulizer::extract_tables(file = pdf, output = "data.frame")
  tab_df <- data.frame()
  df <- func_list2df(tabs) %>%
    dplyr::mutate_at(.vars = 3:ncol(.), ~gsub(",", "", x = .)) %>%
    dplyr::mutate_at(.vars = 3:ncol(.), as.numeric) %>%
    dplyr::mutate(date = gsub('.pdf|_1', '', basename(pdf)),
                  date = paste0('1', date),
                  date = as.Date(date, '%d%b%y')) %>% 
    as.data.frame()
  
  ## check if the total number is correct
  print(colSums(df[3:6], na.rm = T))
  ## assign a unique name to each dataframe
  name <- paste0('df_', gsub('.pdf', '', basename(pdf))); print(name)
  assign(name, df)
}


# Put them in a list
pat <- paste0('^df_.*', yr, '_1$')
ls <- lapply(ls(pattern = pat), get)

# or more directly
ls <- mget(ls(pattern = pat))
dfs <- do.call(rbind, ls)
rownames(dfs) <- NULL

## fix country names
dfs <- dfs %>%
  mutate(Country = gsub('Rep of Macedonia', 'Republic of Macedonia', Country))

dfs_yearly <- dfs %>%
  dplyr::mutate(year = year(date), 
                month = month(date)) %>%
  # dplyr::filter(month == 8) %>%
  group_by(Country, year) %>%
  dplyr::summarise_if(is.numeric, sum, na.rm = T) %>%
  dplyr::select(-month) %>%
  left_join(., ctr_iso, by = c('Country' = 'Contributing_Country')) %>%
  arrange(!is.na(iso3), Country) %>%
  dplyr::select(iso3, Country, everything()) %>%
  as.data.frame()

### `dfs_total` is the same as `dfs_yearly`
# dfs_total <- dfs %>%
#   select(Country, Total, date) %>%
#   pivot_wider(names_from = date, values_from = Total) %>%
#   # mutate(Total_mean = rowMeans(select(., starts_with("2015")), na.rm = TRUE)) %>%
#   mutate(Total_mean = rowSums(select(., starts_with("2015")), na.rm = TRUE)) %>%
#   select(Country, Total_mean) %>%
#   left_join(., ctr_iso, by = c('Country' = 'Contributing_Country')) %>%
#   arrange(!is.na(iso3), Country) %>%
#   dplyr::select(iso3, Country, everything()) %>%
#   as.data.frame()
```




  *test code*
```{r eval=FALSE, include=FALSE}
### --> test code

tabs[1] %>% as.data.frame() %>% select(where(not_all_na)) %>% head(3)#select(!starts_with("X"))
tabs[2] %>% as.data.frame() %>% select(where(not_all_na)) %>% head(3)
tabs[3] %>% as.data.frame() %>% select(where(not_all_na))# %>% head(3)

tab <- tabs[3] %>% as.data.frame() %>%  mutate_all(na_if,"")

## *** in some cases, there is disarrangement in the 1st column! We thus need to 
##  drop the left NAs <in the 1st column> in a dataframe and left shifting the cleaned rows
t1 <- t(apply(tab, 1, function(x) {
  if (is.na(x[1])) {
    y <- x[-seq_len(which.min(is.na(x))-1)]
    length(y) <- length(x)
    y
  } else x
})) %>%
  as.data.frame() %>%
  select(where(not_all_na))


## *** but in some cases, there is disarrangement in the 2nd column too! We thus need to 
##    drop all left NAs <in the 2nd column> in a dataframe and left shifting the cleaned rows. 
## To do so, we first remove the 1st column as it does not have any NAs, 
##    then, we use the same approach as the above
## First, make sure there are non-NA in the 2nd column - I simply check if there are more NAs than 
##    the full length of the dataframe
if (sum(is.na(t1[,2])) > nrow(t1)/2) { 
  
  t2 <- t(apply(t1[,-1], 1, function(x) {
    if (is.na(x[1])) {
      y <- x[-seq_len(which.min(is.na(x))-1)]
      length(y) <- length(x)
      y
    } else x
  })) %>%
    as.data.frame()
  
  t2 <- cbind(t1[,1], t2) ## need to add the 1st column back
    
} else t2 <- t1


tab <- t2 %>% as.data.frame() %>% 
  mutate_all(as.character) %>%
  mutate_all(na_if,"") %>%
  select(where(not_all_na))
names(tab) <- c("No", "Country", "Police", "UNMEM", "Troops", "Total" )



str(tab)
tab
```


## SIPRI data
  
```{r}

xls <- paste0(dirname(dir.peace), '/SIPRI peacekeeping ops/SIPRI data peace ops 2008-18_Yingjie.xlsx')
df2 <- readxl::read_excel(path = xls) %>%
  dplyr::select(-c(month, Date, `Location single`, Region)) %>%
  dplyr::select(Organization, location_code, year, everything()) %>%
  rename(from    = Organization, 
         to      = location_code,
         to_name = location_name) %>%
  # mutate(date = as.Date(Last_Reporting_Date, format = '%m/%d/%Y'),
  #        year = year(date),
  #        total= Female_Personnel + Male_Personnel) %>%
  dplyr::filter(year == 2015) %>%
  as.data.frame()


names(df2)

```



```{r}
# library(textreadr)
# f <- read_rtf(file = 'Trade-Register-2000-2019.rtf', skip = 1)
```

