---
title: "UN_peacekeeping"
author: "Yingjie"
date: "2/23/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---



Troop and Police Contributions
  UN police and military personnel are "contributed" to serve under the blue flag from over 100 countries to carry out the mandates defined by the Security Council. Military operations aim to protect populations against threats and contribute to a secure environment. Policing is central in preventing, mitigating and resolving violent conflict and contributing to early peace-building and reconciliation. Uniformed personnel are contributed and deployed under several different types of arrangements.

# Package and path
```{r }

### To clear your environment 
# remove(list = ls())

## tidy data
library(tidyverse)
library(lubridate)
library(viridis)
# library(xlsx)
library(readxl)
library(readr)

### set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
setwd(dir)
```

## Functions
```{r}
not_any_na <- function(x) all(!is.na(x))
not_all_na <- function(x) any(!is.na(x))
```



# Data

## UN data

### clean _contributions_ data
```{r }
### data
# download from https://peacekeeping.un.org/en/data-troop-and-police-contributions

getwd()
xls <- './UN peacekeeping operations/odp_contributionranks_fixCountryName.xlsx'
ctr_iso  <- readxl::read_excel(path = xls, sheet = 1) %>%
  ungroup() %>%
  distinct(ISOCode3, Contributing_Country)
names(ctr_iso) <- c('iso3', 'Contributing_Country')



# 'BY GENDER' dataset has the most comprehensive vars
csv <- './UN peacekeeping operations/odp_contributionsbygender.csv'
df <- read_csv(file = csv) %>%
  mutate(date = as.Date(Last_Reporting_Date, format = '%m/%d/%Y'),
         year = year(date),
         total= Female_Personnel + Male_Personnel) %>%
  as.data.frame()

str(df)
names(df)
unique(df$Personnel_Type)


### to calculate the total contribution, by each specific date
df_ym <- df %>% as.data.frame() %>%
  dplyr::filter(year == 2015) %>%
  dplyr::select(ISOCode3, 
                Contributing_Country,
                Personnel_Type,
                Mission_Acronym,
                # Female_Personnel, Male_Personnel, 
                Last_Reporting_Date, 
                year,
                total, date) %>%
  # filter(date == as.Date('2015-12-31')) %>%
  group_by(ISOCode3, Contributing_Country, Mission_Acronym, year, date, Personnel_Type) %>%
  summarise(value = sum(total)) %>%
  spread(key = Personnel_Type, value = value) %>%
  dplyr::mutate(total_contr = sum(`Experts on Mission`, Troops, `Individual Police`, `Formed Police Units`, na.rm = T))

### to calculate the total contribution, by year
df_yr <- df_ym %>%
  ungroup() %>%
  dplyr::select(-date) %>%
  group_by(ISOCode3, Contributing_Country, Mission_Acronym, year) %>%
  summarise_all(list( ~ sum(., na.rm = TRUE)))


### to list all the `Mission` names
unique(df_yr$Mission_Acronym)
```



### clean _peacekeeping operations_

  Not sure where this data comes from...
```{r}
### where is the data from? ------------------------------------------------------------------------
getwd()
xls <- './UN peacekeeping operations/UN peacekeeping operations.xlsx'
df  <- readxl::read_excel(path = xls, sheet = 1) 
df_op <- df %>%
  separate(col = 'MISSION NAME', into = c('name1', 'region'), 
           sep = ' in | for ', remove = F) %>%
  dplyr::select(`MISSION NAME`, name1, ACRONYM, Headquarters, Location, region, countries, everything()) %>%
  rename(from      = ACRONYM,
         from_name = name1,
         from_geo  = Headquarters, 
         to        = Location,
         to_region = region)
names(df_op)
# write.xlsx(x = dt, file = xls, sheetName = 'col_separate', row.names = F, append = T)
```


  More accurate data. 
```{r}
###
library(tabulizer)
pdf <- './UN peacekeeping operations/pdf_byMission/dec15_3.pdf'
tabs <- tabulizer::extract_tables(file = pdf, output = "data.frame", guess = T)

tab_df <- data.frame()

for (i in 1:length(tabs)) {
  
  tab <- tabs[i] %>% as.data.frame() %>% select(where(not_all_na)) %>% #select(!starts_with("X"))
    mutate_all(na_if,"") %>% 
    select(where(not_all_na))
  
  print(ncol(tab))
  
  ## We notice the number of columns in each page is different, and only 7 is correct. 
  ## - if ncol = 6, the last column should be separated into two. 
  ## - if ncol = 8, the last two columns should be combined (or added as they are numbers)
  if (ncol(tab) == 6) {
    tab <- tab %>% separate(col = ncol(.), into = c('s1', 's2'), sep = ' ')
    
    } else if (ncol(tab) == 8) {
      
      tab1 <- tab %>% 
        dplyr::mutate(
          a = .[, ncol(.)-1],
          b = .[, ncol(.)-0]) %>%
        dplyr::mutate_at(.vars = c('a', 'b'), ~gsub(",", "", x = .)) %>%
        dplyr::mutate_at(.vars = c('a', 'b'), as.numeric) %>%
        rowwise() %>%
        dplyr::mutate(s = sum(a, b, na.rm = T)) %>%
        select(-a, -b) %>%
        as.data.frame()
      n <- ncol(tab1); n
      tab1 <- tab1[, -c((n-1), (n-2))]
      tab <- tab1
      
      } else {
        tab = tab
      }
  
  print(paste0(' -- i = ', i, ' -- ', ncol(tab)))
  
  ## bind all the pages
  names(tab) <- c('Country', 'UN_Mission', 'Description', 'Male', 'Female', 'Mission', 'Total')
  tab_df <- rbind(tab_df, tab)
}


library(tidyr)

unique(tab_df$UN_Mission)  %>% sort() ## there are some non- `UN_Mission` names were mis-read into this column --> to be removed...
unique(tab_df$Description) %>% sort() ## there are some `Country` names were mis-read into this column 
unique(tab_df$Male) %>% sort()        ## there are some `Country` names were mis-read into this column
unique(tab_df$Female) %>% sort()      ## there are some `Country` names AND `UN_Mission` names were mis-read into this column

tab_df1 <- tab_df %>%
  # filter(!is.na(Country) | !is.na(Mission)) %>%
  mutate(UN_Mission2  = replace(x = UN_Mission,  grepl("Expert|Troop|Individual|Formed", UN_Mission), NA),  ## remove noise in this column
         ctr1 = replace(x = Description, grepl("Expert|Troop|Individual|Formed", Description), NA), ## to get `Country` names
         ctr1 = gsub(',|[[:digit:]]+', '', ctr1), 
         ctr2 = gsub(',|[[:digit:]]+', '', Male),
         ctrM = gsub(',|[[:digit:]]+', '', Female),
                  ) %>%
  mutate_all(na_if,"") %>% 
  dplyr::select(-Male, -Female)%>%
  as.data.frame()


unique(tab_df1$Country)
unique(tab_df1$UN_Mission2)

tab_df2 <- tab_df1 %>%
  ungroup() %>%
  ### --> fix `Country`: remove noise (i.e., any of `UN_Mission`) from `Country` names--------------
  dplyr::mutate(Country2 = ifelse(Country %in% unique(tab_df1$UN_Mission2), NA, Country),
                ctr3     = ifelse(ctrM %in% unique(tab_df1$UN_Mission2), NA, ctrM),
                m        = ifelse(is.na(ctr3), ctrM, NA),
                Country2 = ifelse(is.na(Country2), ctr1, Country2),
                Country2 = ifelse(is.na(Country2), ctr2, Country2),
                Country2 = ifelse(is.na(Country2), ctr3, Country2),
                ) %>%
  dplyr::select(Country, Country2, ctr1, ctr2, ctrM, UN_Mission, UN_Mission2, m, everything()) %>%
  dplyr::select(-Description, -ctr1, -ctr2, -ctrM, -ctr3)%>%
  ### Fill in missing values with previous or next value
  tidyr::fill(Country2) %>%
  
  ### --> fix `UN_Mission2`: get `UN_Mission` names that are in the `Country` column ---------------
  dplyr::mutate(UN_Mission2 = ifelse(is.na(UN_Mission2) & is.na(Country2), Country, UN_Mission2),
                UN_Mission2 = ifelse(is.na(UN_Mission2), m, UN_Mission2)) %>%
  dplyr::select(Country2, UN_Mission, UN_Mission2, m, everything()) %>%
  dplyr::select(-UN_Mission, -Country, -m)%>%
  as.data.frame()
  

  
tab_df3 <- tab_df2 %>%  
  
  ### --> fix `Total`: 1. some of the `Total` are in the `Mission` column -----------------------------
  mutate(Total    = ifelse(is.na(Total), Mission, Total)) %>%
  
  ### some of the `Mission` name and `Total` are in the same cell of `Total`, we should separate them
  mutate(Mission2 = gsub(',|[[:digit:]]+', '', Mission),       ## get the `Mission` name only
         m1       = gsub(',|[[:digit:]]+', '', Total),         ## get the `Mission` name only
         Total2   = as.numeric(gsub("\\D", "", Total)),
         t1       = as.numeric(gsub("\\D", "", Mission))) %>%    ## get the `Total` number only
  
  dplyr::mutate(Mission2 = dplyr::case_when(is.na(Mission2) ~ m1,
                                     Mission2 == m1  ~ Mission2,
                                     T               ~ paste0(Mission2, ' ', m1))) %>%
  as.data.frame() %>%
  ungroup() %>%
  dplyr::mutate(Mission3 = ifelse(is.na(Mission2) & !is.na(UN_Mission2), paste0('999', UN_Mission2), Mission2)) %>%
  # select(Mission, Mission2, m1, everything()) %>%
  select(-m1) %>%
  mutate_all(na_if,"") %>% 
  
  
  ### --> fix `Total`: 2. -----------------------------
  mutate(Total2 = ifelse(is.na(Total2), t1, Total2)) %>%
  
  ### some `Mission` names are in `Female` column, but should exclude values from Female if they are numbers
  # mutate(Mission2 = ifelse(is.na(Mission2), 
  #                          gsub('[[:digit:]]+', '', Female), 
  #                          Mission2)) %>%
  
  # mutate_all(na_if,"") %>% 
  # arrange(is.na(Description2)) %>%
  select(-t1, -Total, -Mission) %>% ## remove the standby data
  as.data.frame()

str(tab_df3)

### clean the data and only keep operation by county data
tab_df4 <- tab_df3 %>%
  dplyr::mutate(
    Country2 = trimws(Country2),
    Mission2 = trimws(Mission2),
    Country2 = stringr::str_squish(Country2),
    Mission2 = stringr::str_squish(Mission2)) %>%             
  dplyr::select(-UN_Mission2, -Mission3) %>%
  dplyr::filter(!is.na(Mission2) & !is.na(Total2)) %>%
  group_by(Country2, Mission2) %>%
  summarise(max = max(Total2, na.rm = T)) %>%
  ungroup() %>%
  # dplyr::mutate(check = Country2 == Mission2) %>%
  dplyr::filter(Country2 != Mission2,
                !Mission2 %in% unique(.$Country2))
  
  

str(tab_df4)
### ??? still, many countries are missing from the data ... 



### test code -------------------------------------------------------------------------------------
# tab1 <- tabs[1] %>% as.data.frame() %>% select(where(not_all_na)) %>% 
#       mutate_all(na_if,"") %>% select(where(not_all_na))
# tab2 <- tabs[2] %>% as.data.frame() %>% select(where(not_all_na)) %>% 
#       mutate_all(na_if,"") %>% select(where(not_all_na))
# tab3 <- tabs[7] %>% as.data.frame() %>% select(where(not_all_na)) %>% 
#       mutate_all(na_if,"") %>% select(where(not_all_na))
# 
# dfs <- do.call(rbind, tabs)
```




### read pdf - contribution 
```{r}
library(tabulizer)
pdf <- './UN peacekeeping operations/pdf/aug15_1.pdf'
tabs <- tabulizer::extract_tables(file = pdf, output = "data.frame", guess = T) 


tab_df <- data.frame()

func_list2df <- function(tabs){
  for (i in 1:length(tabs)) {
    tab <- tabs[i] %>% as.data.frame() %>% select(where(not_all_na)) %>% #select(!starts_with("X"))
      mutate_all(na_if,"") %>% select(where(not_all_na))
    
    ## *** in some cases, there is disarrangement in the 1st column! We thus need to 
    ##  drop the left NAs <in the 1st column> in a dataframe and left shifting the cleaned rows
    t1 <- t(apply(tab, 1, function(x) {
      if (is.na(x[1])) {
        y <- x[-seq_len(which.min(is.na(x))-1)]
        length(y) <- length(x)
        y
      } else x
    })) %>%
      as.data.frame() %>%
      select(where(not_all_na))
    
    
    ## *** but in some cases, there is disarrangement in the 2nd column too! We thus need to 
    ##    drop all left NAs <in the 2nd column> in a dataframe and left shifting the cleaned rows. 
    ## To do so, we first remove the 1st column as it does not have any NAs, 
    ##    then, we use the same approach as the above
    ## First, make sure there are non-NA in the 2nd column - I simply check if there are more NAs than 
    ##    the full length of the dataframe
    if (sum(is.na(t1[,2])) > nrow(t1)/2) { 
      
      t2 <- t(apply(t1[,-1], 1, function(x) {
        if (is.na(x[1])) {
          y <- x[-seq_len(which.min(is.na(x))-1)]
          length(y) <- length(x)
          y
        } else x
      })) %>%
        as.data.frame()
      
      t2 <- cbind(t1[,1], t2) ## need to add the 1st column back
        
    } else t2 <- t1
    
    
    ## clean up the table
    tab <- t2 %>% as.data.frame() %>% 
      mutate_all(as.character) %>%
      mutate_all(na_if,"") %>%
      select(where(not_all_na))
    names(tab) <- c("No", "Country", "Police", "UNMEM", "Troops", "Total" )
    
    ## bind all the pages
    tab_df <- rbind(tab_df, tab)
  }
  return(tab_df)
}


### test function 
df <- func_list2df(tabs) %>%
  # mutate_all(~(gsub(",", "", .))) %>% ## this also works
  dplyr::mutate_at(.vars = 3:ncol(.), ~gsub(",", "", x = .)) %>%
  dplyr::mutate_at(.vars = 3:ncol(.), as.numeric) %>%
  as.data.frame()

str(df)

colSums(df[3:6], na.rm = T)




pdfs <- list.files(path = './UN peacekeeping operations/pdf/', pattern = '*.pdf$', full.names = T); pdfs
for (pdf in pdfs) {
  tabs <- tabulizer::extract_tables(file = pdf, output = "data.frame")
  tab_df <- data.frame()
  df <- func_list2df(tabs) %>%
    dplyr::mutate_at(.vars = 3:ncol(.), ~gsub(",", "", x = .)) %>%
    dplyr::mutate_at(.vars = 3:ncol(.), as.numeric) %>%
    dplyr::mutate(date = gsub('.pdf|_1', '', basename(pdf)),
                  date = paste0('1', date),
                  date = as.Date(date, '%d%b%y')) %>% 
    as.data.frame()
  
  ## check if the total number is correct
  print(colSums(df[3:6], na.rm = T))
  ## assign a unique name to each dataframe
  name <- paste0('df_', gsub('.pdf', '', basename(pdf))); print(name)
  assign(name, df)
}


# Put them in a list
ls <- lapply(ls(pattern = "^df_.*15_1$"), get)

# or more directly
ls <- mget(ls(pattern = "^df_.*15_1$"))
dfs <- do.call(rbind, ls)
rownames(dfs) <- NULL

dfs_total <- dfs %>%
  select(Country, Total, date) %>%
  mutate(Country = gsub('Rep of Macedonia', 'Republic of Macedonia', Country)) %>%
  pivot_wider(names_from = date, values_from = Total) %>%
  mutate(Total_mean = rowMeans(select(., starts_with("2015")), na.rm = TRUE)) %>%
  select(Country, Total_mean) %>%
  left_join(., ctr_iso, by = c('Country' = 'Contributing_Country')) %>%
  arrange(!is.na(iso3), Country) %>%
  as.data.frame()
```





```{r test code, eval=FALSE, include=FALSE}
### --> test code

tabs[1] %>% as.data.frame() %>% select(where(not_all_na)) %>% head(3)#select(!starts_with("X"))
tabs[2] %>% as.data.frame() %>% select(where(not_all_na)) %>% head(3)
tabs[3] %>% as.data.frame() %>% select(where(not_all_na))# %>% head(3)

tab <- tabs[3] %>% as.data.frame() %>%  mutate_all(na_if,"")

## *** in some cases, there is disarrangement in the 1st column! We thus need to 
##  drop the left NAs <in the 1st column> in a dataframe and left shifting the cleaned rows
t1 <- t(apply(tab, 1, function(x) {
  if (is.na(x[1])) {
    y <- x[-seq_len(which.min(is.na(x))-1)]
    length(y) <- length(x)
    y
  } else x
})) %>%
  as.data.frame() %>%
  select(where(not_all_na))


## *** but in some cases, there is disarrangement in the 2nd column too! We thus need to 
##    drop all left NAs <in the 2nd column> in a dataframe and left shifting the cleaned rows. 
## To do so, we first remove the 1st column as it does not have any NAs, 
##    then, we use the same approach as the above
## First, make sure there are non-NA in the 2nd column - I simply check if there are more NAs than 
##    the full length of the dataframe
if (sum(is.na(t1[,2])) > nrow(t1)/2) { 
  
  t2 <- t(apply(t1[,-1], 1, function(x) {
    if (is.na(x[1])) {
      y <- x[-seq_len(which.min(is.na(x))-1)]
      length(y) <- length(x)
      y
    } else x
  })) %>%
    as.data.frame()
  
  t2 <- cbind(t1[,1], t2) ## need to add the 1st column back
    
} else t2 <- t1


tab <- t2 %>% as.data.frame() %>% 
  mutate_all(as.character) %>%
  mutate_all(na_if,"") %>%
  select(where(not_all_na))
names(tab) <- c("No", "Country", "Police", "UNMEM", "Troops", "Total" )



str(tab)
tab
```


## SIPRI data
  
```{r}
# \Data\data_01_raw\Arm trade_Peacekeeping\SIPRI peacekeeping ops\SIPRI data peace ops 2008-18_Yingjie.xlsx


xls <- './SIPRI peacekeeping ops/SIPRI data peace ops 2008-18_Yingjie.xlsx'
df2 <- readxl::read_excel(path = xls) %>%
  dplyr::select(-c(month, Date, `Location single`, Region)) %>%
  dplyr::select(Organization, location_code, year, everything()) %>%
  rename(from    = Organization, 
         to      = location_code,
         to_name = location_name) %>%
  # mutate(date = as.Date(Last_Reporting_Date, format = '%m/%d/%Y'),
  #        year = year(date),
  #        total= Female_Personnel + Male_Personnel) %>%
  dplyr::filter(year == 2015) %>%
  as.data.frame()


names(df2)

```



```{r}
library(textreadr)
f <- read_rtf(file = 'Trade-Register-2000-2019.rtf', skip = 1)
```

