---
title: "Untitled"
author: "Yingjie"
date: "11/21/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---


#  Dirs and Packages

```{r}
# To clear your environment 
# remove(list = ls())

# set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
setwd(dir)
getwd()

dir_results <- './AidData_SDGs/_results_aiddata_sdgs/'

today <- format(Sys.time(), "%Y%m%d"); print(today)

library(cowplot)
library(scales)    ## percent 

# install.packages("devtools")
library(devtools)
# install_github("mdilorenzo/aidtools")
library(aidtools)
library(rgdal)
library(rgeos)
library(foreach)
library(tm)
library(tidyverse)
# library(xlsx)
```


# 1. Data
```{r}
## read data
dir.sdgv1 <- './AidData_SDGs/Financing_to_the_Sustainable_Development_Goals_Dataset_version_1_0/'
csv.sdgv1 <- paste0(dir.sdgv1, 'FinancingtotheSDGsDataset_v1.0.csv')
# sdgv1 <- read.csv(csv.sdgv1, stringsAsFactors = F)
# save(sdgv1, file = './FinancingtotheSDGsDataset_v1.0.RData') ### save as RData
load(file = './FinancingtotheSDGsDataset_v1.0.RData')        ### load the RData

## take a subset
sdgv1.sub <- sdgv1[1:100, ]
str(sdgv1.sub)
names(sdgv1.sub)
```



## 1.1 Data discription
```{r}
# Data discription --------------------------------------------------------
df <- sdgv1

### how many years covered in the dataset: 2000 - 2013
yrs <- sort(unique(df$year)); print(yrs)

### how many countries in the dataset - (78 donors)
donors <- df %>%
  distinct(donor); print(head(donors))

### how many countries in the dataset - (214 receients)
recipients <- df %>%
  distinct(recipient); print(head(recipients))

# write.xlsx(donors,     './aiddata_sdgs_country_list.xlsx', 
#            row.names = F, sheetName = 'donors', append = T)
# write.xlsx(recipients, './aiddata_sdgs_country_list.xlsx', 
#            row.names = F, sheetName = 'recipients', append = T)

### how much aids were distributed to each goal, by year?
str(sdgv1)
aid_year_goal <- sdgv1 %>%
  group_by(year) %>%
  summarise_at(vars(goal_1:goal_17), mean, na.rm = TRUE) %>%
  gather(key = 'goal', value = 'value', goal_1:goal_17)

aid_year_goal$goal <- 
  factor(aid_year_goal$goal, levels = paste0('goal_', seq(1,17)))

hist(aid_year_goal$value)



library(ggplot2)
library(hrbrthemes)
library(RColorBrewer)
library(viridis)


library(classInt)
library(RColorBrewer)
nclass = 10
classint <- classIntervals(aid_year_goal$value, n = nclass, style = "quantile") #can use "jenks" but is very slow
brks <- as.integer(classint$brk); brks #approximate quantiles
labs <- brks[-1]; labs
aid_year_goal$breaks <- cut(aid_year_goal$value, breaks=brks, labels=labs)


# myPalette <- colorRampPalette(rev(brewer.pal(11, "RdYlBu")));# myPalette
# sc <- scale_fill_gradientn(colours = myPalette(100)) # , limits=c(1, 8)


ggplot(aid_year_goal, aes(x=year, y=goal, 
                          # fill= (breaks), 
                          fill = (value))) + 
  geom_tile() +
  # scale_fill_gradient(low="red", high="blue") + 
  # scale_fill_distiller(palette = "RdPu") +
  scale_fill_viridis(discrete=FALSE, direction = -1, labels = comma) +
  # scale_fill_brewer('Legend name', palette = "YlGnBu") +
  # scale_fill_brewer('Legend name', palette = 'BrBG') +
  # sc+
  
  # guides(fill = guide_legend(label.hjust = 1, label = T))+
  labs(fill = 'Aids ($)') +
  # scale_fill_continuous(labels = scales::unit_format(unit = "", scale = 10))
  theme_minimal() +
  xlab('') + ylab('')

getwd()
ggsave(filename = './AidData_SDGs/_results_aiddata_sdgs/aids_to_goals_overTime.png', 
       plot = last_plot(), width = 6, height = 4, units = 'in', dpi = 300)
# data <- sdgv1 %>%
#   group_by(year) %>%
#   summarise_at(vars(goal_1:goal_17), mean, na.rm = TRUE) %>% as.matrix() %>% t()
# # Default Heatmap
# heatmap(data)
# # No dendrogram nor reordering for neither column or row
# heatmap(data, Colv = NA, Rowv = NA, scale="row", color = )
# legend(x="left", legend=c("min", "ave", "max"), 
#      fill=colorRampPalette(brewer.pal(8, "Oranges"))(3))
```

  We need to make sure the country names are consistant with all the data we are using.
  I will tidy and match this manually in the xlsx files. 



## 1.2 Map the flow
```{r}
## take a subset
names(df)
df13 <- df %>% filter(year >= 2013) %>%
  select(donor, recipient, goal_13) %>%
  group_by(donor, recipient) %>% ## the data is project-by-project, so we need to combine them together
  dplyr::summarise(goal_i = sum(goal_13, na.rm = T)) %>%
  spread(key = recipient, value = goal_i) ## make a matrix
```



### Sankey diagram
```{r}
# Load package
library(networkD3)
library(tidyverse)

### dataframe format
' 
---------------------------------
       |  A  |  B  |  C  | ...
---------------------------------
A      |  1  |  2  |  5  | ...
---------------------------------
B      | 4   |  7  |  9  |
---------------------------------
...    | ...       | ...
---------------------------------
'

data <- df13 %>% as.data.frame()
row.names(data) <-data[,1] ## 1st column as row name
data <- data[,-1]          ## remove the 1st  column

# Transform it to connection data frame with tidyr from the tidyverse
links <- data %>% 
  as.data.frame() %>% 
  rownames_to_column(var="source") %>% 
  gather(key="target", value="value", -1) %>%
  filter(value != 0) %>%
  # filter(value >= 10^6) %>%
  top_n(50)
hist(links$value, breaks = 10000, xlim = c(0, 10^7))

# From these flows, we need to have a "node" dataframe, which lists all the entities involved in the flow
nodes <- data.frame(name=c(as.character(links$source), as.character(links$target)) %>% unique())

names(data)
# write.csv(links, 'links.csv')


## With networkD3, connection must be provided using id, but not the real name like in the "links" dataframe. So we need to reformat it.
links$IDsource=match(links$source, nodes$name)-1 
links$IDtarget=match(links$target, nodes$name)-1


####################################################
# Add a 'group' column to each connection:
links$group = as.factor(row.names(links)) 


# Add a 'group' column to each node. Here I decide to put all of them in the same group to make them grey
nodes$group=as.factor(nodes$name)


# prepare color scale: I give one specific color for each node.
# Give a color for each group:

# my_color <- 'd3.scaleOrdinal().domain(["Forest_1999", "Water_1999", "BareGround_1999",
#                                        "Forest_2013", "Water_2013", "BareGround_2013",
#                                        "Forest_2018", "Water_2018", "BareGround_2018", 
#                                        "f","w","b"]).range(["green", "blue", "orange", 
#                                                             "green", "blue", "orange", 
#                                                             "green", "blue", "orange", 
#                                                             "lightgreen", "#6baed6", "yellow"])'
# 

# Make the Network
sankeyNetwork(Links = links, Nodes = nodes, 
              Source = "IDsource", Target = "IDtarget", 
              Value = "value", NodeID = "name", 
              # colourScale=my_color, 
              fontSize=12,
              LinkGroup="group", 
              NodeGroup="group")

```








##############################################################################
# 2. Cal aids allocation at target level
##############################################################################
```{r}
# cal aids allocation at target level -------------------------------------
old_sdg_dataset <- sdgv1.sub

new_target_coded_data <- target_coder(dat = old_sdg_dataset, coalesced_purpose = T)
new_target_coded_data <- left_join(old_sdg_dataset[, 1:11], new_target_coded_data, by = "aiddata_id")

## NOT WORK
new_goal_coded_data   <- sdg_coder(dat = old_sdg_dataset, coalesced_purpose = T)

# load('./aidtools-master/data/sdgs.rda')
# load('./aidtools-master/data/combined-target-weights.rda')
```






## 2.1 Explore data
```{r}
names(aid)
str(aid)
summary(aid)
## years: 2000 - 2014
unique(aid$year) %>% sort()
## countries
unique(aid$donors_iso3) %>% sort()     ## only China
unique(aid$recipients) %>% sort()      ## 168 recipients
unique(aid$all_recipients) %>% sort()  ## 168
unique(aid$recipients_iso3) %>% sort() ## 172
unique(aid$recipient_iso3) %>% sort()  ## 174
ctr <- aid %>% distinct(recipients, recipients_iso3, recipient_iso3, all_recipients)
getwd()
file_ctr <- paste0(dir_results, 'recipients_in_chinese_aiddata.xlsx'); file_ctr
# write.xlsx(x = ctr, file = file_ctr, sheetName = 'recipients', append = T, row.names = F)


## sectors
unique(aid$ad_sector_names) %>% sort() ## 24 secors
unique(aid$crs_sector_name)            ## 24 secors
sectors <- aid %>% distinct(ad_sector_names, crs_sector_name)


unique(aid$ad_purpose_names) %>% sort() ## NA
unique(aid$flow)
unique(aid$loan_type)
unique(aid$gazetteer_adm_name)
```


## 2.2 Aggregate data
```{r}
## total aids by sectors
aid_type <- aid %>% 
  dplyr::filter(!is.na(usd_current) & !is.na(ad_sector_names)) %>%
  dplyr::group_by(ad_sector_codes, ad_sector_names) %>%
  dplyr::summarise(usd_sect = sum(usd_current, na.rm = T)) %>%
  ungroup() %>%
  dplyr::mutate(usd_tot = sum(usd_sect)) %>%
  dplyr::mutate(sect_pct = usd_sect / usd_tot) # percent

aid_type$ad_sector_names

## aid for what? - transport 1/6; health; etc.
ggplot(data = aid, aes(x=ad_sector_names)) + 
  geom_bar() + 
  coord_flip() + 
  theme_bw()

ggplot(data = aid_type, aes(x = 1, y = sect_pct, 
                            fill = ad_sector_names)) + 
  geom_bar(width = 1, stat = 'identity') + 
  coord_polar("y", start=0, direction = -1)+
  theme_void()+
  geom_text(aes(label = paste0(#ad_sector_names, '\n', 
                               scales::percent(round(sect_pct, digits = 3)))),
            position = position_stack(vjust = 0.5))


ggplot(data = aid_type, aes(x = reorder(ad_sector_names, sect_pct), 
                            y = sect_pct)) + 
  geom_col() + 
  coord_flip() + 
  scale_y_continuous(labels = scales::percent) +
  ylab('Percentage of the total') + xlab('Aid sector names') +
  theme_bw()
```


## 2.3 Subset data
```{r}
aid_type <- aid_type %>% 
  arrange(desc(usd_sect))
head(aid_type)

top5 <- as.vector((aid_type$ad_sector_names)[1:5]); print(top5)

aid_type_top5 <- aid_type %>% dplyr::filter(ad_sector_names %in% top5)
```




# 4. Map aids

```{r}
nrow(aid)
unique(aid$longitude) %>% length()

names(aid)
str(aid$usd_current)

library(sf)
library(maps)
library(mapdata)
```


##  Total
### by coords
```{r}
aid_agg <- aid %>% 
  dplyr::filter(!is.na(usd_current) & !is.na(latitude) & !is.na(longitude)) %>%
  dplyr::group_by(#project_id, 
           # project_location_id, place_name,	, 
           # recipients, 
           # ad_sector_names, 
           latitude,	longitude) %>%
  dplyr::summarise(usd_tot = sum(usd_current, na.rm = T))
nrow(aid_agg)

shp <- map_data('world')
geo <- aid_agg

head(shp)
head(geo)

hist(geo$usd_tot)

ggplot() + 
  geom_polygon(data = shp, aes(x=long, y = lat, group = group), fill = 'gray90') + 
  geom_point(data = geo,  # use geom_point, instead of geom_sf
             aes(x = longitude, 
                 y = latitude, 
                 # size  = log10(usd_tot),
                 color = log10(usd_tot)),
             size = 1,
             alpha = 0.6) +              ## transparancy
  theme_bw() +
  # theme_dark() +
  theme(legend.position = c(0.1, 0.35)) +
  # labs(size = 'Population') +          ## change 
  xlab('') + ylab('') +
  scale_color_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 6, na.value = 'black')
  # scale_color_brewer(type = 'seq', palette = 'Spectral')
  # scale_color_viridis_c(option = "viridis", alpha = 0.6)
  # scale_color_manual('sentiments', values=pal)
  # theme(legend.position = "none")

getwd()
pname <- paste0(dir_results, '/map_aid_pt_', today, '.png'); pname
ggsave(filename = pname, plot = last_plot(), width = 7, height = 4, units = 'in', dpi = 400)

```



### by nation

```{r}
aid_nation <- aid %>% 
  dplyr::filter(!is.na(usd_current) & !is.na(recipients)) %>%
  dplyr::mutate(recipients_iso3 = substr(x = recipients_iso3, start = 1, stop = 3)) %>%
  dplyr::group_by(recipients, recipients_iso3) %>% ## , ad_sector_names
  dplyr::summarise(usd_tot = sum(usd_current, na.rm = T))
  
### read in shp data
library(sf)
dir.shp <- "G:/My Drive/_data/shp/world_shp/ne_10m_admin_0_countries"
shp <- st_read(dsn = dir.shp, layer = "ne_10m_admin_0_countries") %>%
  select(ADMIN, ISO_A3, ISO_A3_EH, ISO_N3)

head(shp)

# ggplot() + 
#   geom_sf(data = shp,
#           aes(fill = ISO_A3),    ## set color for each unit
#           size = 0.01,          ## bordor size
#           show.legend =  T, 
#           color = 'gray'        ## border color
#   ) +      
#   # color_scale +
#   theme_bw() + labs(fill = 'SDGc score') +        ## change legend title
#   theme(legend.position = "none") + ylab('') + xlab('')

geo <- aid_nation
geo <- merge(x = shp, y = geo, by.x = 'ISO_A3', by.y = 'recipients_iso3', all.x = T)

str(shp)
head(shp)
head(geo)
nrow(geo)

hist(geo$usd_tot)

### label - Display log tick marks
brks1 <- sort(log10(unique(geo$usd_tot))); brks1
min  <- round(min(log10(unique(geo$usd_tot)), na.rm=TRUE)); min
max  <- round(max(log10(unique(geo$usd_tot)), na.rm=TRUE)); max

brks <- seq(min, max, 1); brks
labs <- parse(text=paste0("10^", brks)); labs

## ref: https://stackoverflow.com/questions/10628547/use-superscripts-in-r-axis-labels
## http://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations#display-log-tick-marks

ggplot() + 
  geom_sf(data = geo,
          aes(fill = log10(usd_tot)),   ## set color for each unit
          size = 0.1,                   ## bordor size
          show.legend =  T,
          color = 'gray20')+            ## border color   
  theme_bw() +
  theme(legend.position = c(0.1, 0.35)) +
  scale_fill_gradient2(name = "USD", 
                       low  = 'blue', mid = 'white', high = 'red', 
                       midpoint = 8, na.value = 'gray80', 
                       breaks = brks, labels = labs)
  # facet_wrap(~ad_sector_names) + 
  # labs(size = 'Population') +          ## change 
  # xlab('') + ylab('') +
  # scale_fill_gradient(name   = "US $", trans = "log10", 
  #                     breaks = brks, labels = labs)

getwd()
pname <- paste0(dir_results, '/map_aid_nation_', today, '.png'); pname
ggsave(filename = pname, plot = last_plot(), width = 7, height = 4, units = 'in', dpi = 400)
```



##  Sector
### by sector by nation

```{r}
aid_nation <- aid %>% 
  dplyr::filter(!is.na(usd_current) & !is.na(recipients)) %>%
  dplyr::filter(ad_sector_names %in% top5) %>%
  dplyr::mutate(recipients_iso3 = substr(x = recipients_iso3, start = 1, stop = 3)) %>%
  dplyr::group_by(recipients, recipients_iso3, ad_sector_names) %>% ## , ad_sector_names
  dplyr::summarise(usd_tot = sum(usd_current, na.rm = T)) %>%
  dplyr::filter(!is.na(usd_tot))
  

aid_nation$ad_sector_names <- factor(aid_nation$ad_sector_names, 
                                     levels = top5)

### read in shp data
library(sf)
dir.shp <- "G:/My Drive/_data/shp/world_shp/ne_10m_admin_0_countries"
shp <- st_read(dsn = dir.shp, layer = "ne_10m_admin_0_countries") %>%
  select(ADMIN, ISO_A3, ISO_A3_EH, ISO_N3)

head(shp)

# ggplot() + 
#   geom_sf(data = shp,
#           aes(fill = ISO_A3),    ## set color for each unit
#           size = 0.01,          ## bordor size
#           show.legend =  T, 
#           color = 'gray'        ## border color
#   ) +      
#   # color_scale +
#   theme_bw() + labs(fill = 'SDGc score') +        ## change legend title
#   theme(legend.position = "none") + ylab('') + xlab('')

geo <- aid_nation
geo <- merge(x = shp, y = geo, by.x = 'ISO_A3', by.y = 'recipients_iso3', all.y = T)

str(shp)
head(shp)
head(geo)
nrow(geo)

hist(geo$usd_tot)

### label - Display log tick marks
brks1 <- sort(log10(unique(geo$usd_tot))); brks1
min  <- round(min(log10(unique(geo$usd_tot)), na.rm=TRUE)); min
max  <- round(max(log10(unique(geo$usd_tot)), na.rm=TRUE)); max

brks <- seq(min, max, 1); brks
labs <- parse(text=paste0("10^", brks)); labs

## ref: https://stackoverflow.com/questions/10628547/use-superscripts-in-r-axis-labels
## http://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations#display-log-tick-marks

ggplot() + 
  geom_sf(data = shp,
          aes(fill = NULL),   ## set color for each unit
          size = 0.1,                   ## bordor size
          show.legend =  F,
          color = 'gray30')+   
  geom_sf(data = geo,
          aes(fill = log10(usd_tot)),   ## set color for each unit
          size = 0.01,                   ## bordor size
          show.legend =  T,
          color = 'gray70')+            ## border color   
  
  scale_fill_gradient2(name = "USD", 
                       low  = 'blue', mid = 'white', high = 'red', 
                       midpoint = 8, na.value = 'gray80', 
                       breaks = brks, labels = labs) +
  facet_wrap(~ad_sector_names, ncol = 2) +
  theme_bw() +
  theme(legend.position = c(0.8, 0.15), legend.direction = "horizontal",
        legend.key.size = unit(0.5, "in"))
  # labs(size = 'Population') +          ## change 
  # xlab('') + ylab('') +
  # scale_fill_gradient(name   = "US $", trans = "log10", 
  #                     breaks = brks, labels = labs)

getwd()
pname <- paste0(dir_results, '/map_aid_nation_bySector_', today, '.png'); pname
ggsave(filename = pname, plot = last_plot(), width = 7, height = 7, units = 'in', dpi = 300)
```



## Selected sectors
210: Transport and Storage - 58.8%
230: Energy Generation and Supply - 21.9%

```{r}
aid_infr <- aid %>% 
  dplyr::filter(!is.na(usd_current) & !is.na(latitude) & !is.na(longitude)) %>%
  dplyr::filter(ad_sector_codes %in% c(210, 230)) 

hist(aid_infr$year)

library(dplyr)
aid_infr_yr <- aid_infr %>%
  add_count(year) %>%
  # group_by(year) %>%
  # count(year) #%>%
  # dplyr::rename(year = n.x, n = v.freq)
  ggplot()+
  geom_col(aes(x = year, y = n))

aid_infr_yr
```


