---
title: "FDI"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---


# Set up
```{r Paths and packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
remove(list = ls())

### set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
getwd()

### data path 
source('./Code/_path of data.R')


### packages
source('./Code/_package list.R')
source('./Code/function_scatter_plot.R')

```


# Data

## Ancillary data
```{r}

### extended country-iso3 pair list for matching
load('./Data/Ancillary_Data_iso_eora_ex.RData')  ## ## iso_eora_ex


### use the Eora data as the template 
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')
load(f) # `temp_eora`
```



## Refugee Data Finder - UNHCR

  The UN Refugee Agency, UNHCR, collates population data relating to persons who are forcibly displaced or stateless. The data is sourced primarily from governments and also from UNHCR operations.
  
  People that UNHCR protects and/or assists include those 
  
  * who have been *forcibly displaced* (refugees, asylum-seekers, other people in need of international protection and internally displaced people); 
  * those who are *stateless* (most of whom are not forcibly displaced); 
  * and other groups of concern to whom UNHCR has extended its protection or provided assistance on a humanitarian basis. 
  
  Refugee and IDP returnees are included in the solutions datasets. 

  https://www.unhcr.org/refugee-statistics/methodology
  
  https://www.unhcr.org/refugee-statistics/methodology/data-content (*metadata*)

```{r - preprocessing}
### get the indicator name
ind <- 'Refugee'

library(countrycode)

yrs <- c(2000, 2005, 2010, 2015)

f <- paste0(dir.refugee, 'query_data_1990_2020/data.csv'); f

d <- readr::read_csv(file = f, show_col_types = F, skip = 14) 
names(d)

names(d) <- c('year', 'o_name', 'd_name', 'refugees', 'asylum_seekers', 'returned_refugees', 
              'idps', 'returned_idps', 'stateless', 'ooc', 'oip', 'hst')


df <- d %>%
  select(1:`refugees`) 
names(df)

```


  Country/territory of asylum and origin – the full list of *UN countries* is available on the UNSD methodology page. 
  ISO3 country codes are included. UNHCR uses the following non-standard ISO3 country codes:
  - UKN for Various/unknown
  - STA for Stateless

```{r - preprocessing - country code}


## convert country code using https://github.com/vincentarelbundock/countrycode
ctr_code <- df %>%
  distinct(o_name, d_name) %>%
  pivot_longer(names_to = 'name', values_to = 'ctr_name', cols = 1:2) %>%
  distinct(ctr_name) %>%
  as.data.frame() %>%
  dplyr::mutate(
    # ctr_name = str_to_upper(ctr_name),
    iso = countrycode(
      sourcevar = ctr_name, 
      origin = 'country.name', 
      destination = 'iso3c')) %>%
  dplyr::mutate(
    iso = case_when(
      ctr_name == 'Various' ~ 'UKN',
      ctr_name == 'Unknown' ~ 'UKN',
      ctr_name == 'Stateless' ~ 'STA',
      ctr_name == 'Tibetan' ~ 'TBT', ## "CHN" would cause duplicated iso3c 
      T ~ iso) ) %>%
  arrange(!is.na(iso))

```


```{r - final 📁}

dfl_ <- df %>%
  left_join(x = .,
            y = ctr_code,
            by=c('o_name' = 'ctr_name')) %>%
  dplyr::rename("o_iso" = "iso") %>%
  select(1:o_name, o_iso, everything()) %>%
  left_join(x = .,
            y = ctr_code,
            by=c('d_name' = 'ctr_name')) %>%
  dplyr::rename("d_iso" = "iso") %>%
  select(1:d_name, d_iso, everything())



## format as the same as EORA data
df_m <- dfl_ %>%
  ## put year as column and fill na if needed
  pivot_wider(names_from = 'year', values_from = 'refugees') 



# Target years to fill
target_years <- yrs

# Function to fill NA values in target year columns
fill_na_adjacent_cols <- function(df, target_years) {
  for (year in target_years) {
    year <- as.character(year)  # Ensure year is a string
    
    # Find the previous and next year columns
    year_next <- as.character(as.numeric(year) + 1)
    year_prev <- as.character(as.numeric(year) - 1)
    
    # Check if these columns exist in the data
    if (year_next %in% colnames(df)) {
      df[[year]][is.na(df[[year]])] <- df[[year_next]][is.na(df[[year]])]
    }
    if (year_prev %in% colnames(df)) {
      df[[year]][is.na(df[[year]])] <- df[[year_prev]][is.na(df[[year]])]
    }
  }
  return(df)
}

# Apply the function to the data frame
df_m_filled <- fill_na_adjacent_cols(df_m, target_years) %>%
  dplyr::select(-c(o_name, d_name)) %>%
  pivot_longer(cols = `1990`:ncol(.), names_to = 'year', values_to = 'value') %>%
  dplyr::mutate(year = as.numeric(year)) %>%
  dplyr::filter(year %in% yrs) %>%
  ## the template can limit the years included
  merge(x = temp_eora,
        y = .,
        by.x = c('year', 'iso3', 'to'),
        by.y = c('year', 'o_iso', 'd_iso'), 
        all.x = T) %>%
  as.data.frame()



### To generate the same format as Eora in Matrix
df_m_formatted <- df_m_filled %>%
  ungroup() %>%
  arrange(year, iso3, to) %>%
  dplyr::select(year, ctr, everything()) %>%
  spread(key = to, value = value) %>%
  arrange(year, iso3)

df_m_formatted %>%
  group_by(year) %>% tally()



### save to the `eora_cleaned` folder ----------------------------------------------------
fname <- paste0(dir.eora_cleaned, ind, '_matrix.xlsx'); fname
writexl::write_xlsx(x = df_m_formatted, path = fname)
```





