---
title: "FDI"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---


# Set up
```{r Paths and packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
remove(list = ls())

### set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
getwd()

### data path 
source('./Code/_path of data.R')


### packages
source('./Code/_package list.R')
source('./Code/function_scatter_plot.R')

```


# Data

## Ancillary data
```{r}

### extended country-iso3 pair list for matching
load('./Data/Ancillary_Data_iso_eora_ex.RData')  ## ## iso_eora_ex


### use the Eora data as the template 
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')
load(f) # `temp_eora`
```



## High-skilled immigrants

  [Database on Immigrants in OECD and non-OECD Countries](https://www.oecd.org/en/data/datasets/database-on-immigrants-in-oecd-and-non-oecd-countries.html)
  The datasets include information on demographic characteristics (age and gender), duration of stay, labour market outcomes (labour market status, occupations, sectors of activity), fields of study, educational attainment and the place of birth.
  * DIOC: Contains data on 34 destination countries and more than 200 countries of origin.
  * DIOC-E (DIOC extended): Contains data on 100 destination countries and more than 200 countries of origin. 
  
  **Data range**: 2000, 2010
  
  **Variables**: 
  `country` country of residence ISO 3166-1 (alpha 3)
  `coub` country of birth ISO 3166-1 (alpha 3)
  `regionb` region of birth 
      AFRI = Africa
      ASIA = Asia
      EURO = Europe
      NOAM = Northern America
      OCEA = Oceania
      SCAC = South and Central America and the Caribbean
      UNK = Unknown
  `lfs` labour force status 
      1 = employed
      2 = unemployed
      3 = inactive
      99 = Unknown
  `edu_lfs` or `education` educational attainment (broad) 
      1 = ISCED 0/1/2       low
      2 = ISCED 3/4         medium
      3 = ISCED 5/6/7/8     high
      99 = Unknown
      
  
  *High-skilled workers*, our main focus, are defined as those with at least one year of tertiary education. (Kerr et al 2016)
  
  Kerr, S.P., Kerr, W., Özden, Ç., Parsons, C., 2016. Global Talent Flows. Journal of Economic Perspectives 30, 83–106. https://doi.org/10.1257/jep.30.4.83


```{r - OECD data - preprocessing}

### get the indicator name
ind <- 'High-skilled immigrants'

library(countrycode)

yrs <- c(2000, 2005, 2010, 2015)


## data in 2000 is not available 
# f <- paste0(dir.dico, 'DIOC-E-2000-01/File-2-DIOC-E-2010-11-Labour-Force-Status.csv'); f
# d1 <- readr::read_csv(file = f, show_col_types = F)

f <- paste0(dir.dico, 'DIOC-E-2010-11/File-2-DIOC-E-2010-11-Labour-Force-Status.csv'); f
df <- readr::read_csv(file = f, show_col_types = F) %>%
  dplyr::rename(
    o_iso = coub, 
    d_iso = country) %>%
  mutate(year = 2010) %>%
  as.data.frame()
  
names(df)
unique(df$education)
unique(df$lfs_status)

## aggregate data 
df_1 <- df %>%
  group_by_at(vars(-age, -sex, -number, -reg_region, -fborn, -lfs_status)) %>%
  dplyr::summarise(n = sum(number, na.rm = T)) %>%
  ungroup() %>%
  filter(education == 3) %>%
  select(-education) %>%
  select(o_iso, d_iso, everything()) %>%
  as.data.frame()

```


```{r - WB data}
##' it might be better not to use this data 
##' because it lacks information on education/skill levels
dir <- './Data/data_01_raw/WB_Global Bilateral Migration/'
f <- paste0(dir, 'P_Data_Extract_From_Global_Bilateral_Migration.xlsx'); f
d <- readxl::read_excel(f) 

names(d) 
names(d) <- c("o_name", "o_iso", "Gender", "Gender_Code", 
              "d_name", "d_iso", "1960", "1970", "1980", "1990", "2000")

unique(d$Gender)
## aggregate data 
df_2wb <- d %>%
  filter(Gender == 'Total') %>%
  select(-Gender, -Gender_Code, -o_name, -d_name) %>%
  ungroup() %>%
  gather(3:ncol(.), key = 'year', value = 'n') %>%
  as.data.frame()
```



```{r - Combine data}

## combine two data sets
df_12 <- 
  rbind(df_1, df_2wb) %>% ## WB data do NOT have education/skill levels information
  mutate(n = as.numeric(n)) %>%
  spread(key = year, value = n) %>%
  as.data.frame()


## plot to check data
df_12 %>%
  ggplot() +
  geom_point(aes(x = `2000`, y = `2010`), alpha = 0.3) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  geom_abline(slope = 1, color = 'red') +
  theme_classic()



## update ------------------------------------------------------------------------------
df_ <- 
  # rbind(df_1, df_2) %>% ## WB data do NOT have education/skill levels information  
  df_1 %>%
  mutate(n = as.numeric(n)) %>%
  spread(key = year, value = n) %>%
  as.data.frame()
```



```{r - final 📁}

## format as the same as EORA data
df_m <- df_ %>%
  ## fill data gap using adjacent years
  dplyr::mutate(
    `2005` = NA,
    `2015` = `2010`
  )



# Apply the function to the data frame
df_m_filled <- df_m %>%
  as.data.frame() %>%
  gather(3:ncol(.), key = 'year', value = 'value') %>%
  dplyr::mutate(year = as.numeric(year)) %>%
  dplyr::filter(year %in% yrs) %>%
  ## the template can limit the years included
  merge(x = temp_eora,
        y = .,
        by.x = c('year', 'iso3', 'to'),
        by.y = c('year', 'o_iso', 'd_iso'), all.x = T) %>%
  as.data.frame()



### To generate the same format as Eora in Matrix
df_m_formatted <- df_m_filled %>%
  ungroup() %>%
  arrange(year, iso3, to) %>%
  dplyr::select(year, ctr, everything()) %>%
  spread(key = to, value = value) %>%
  arrange(year, iso3)

df_m_formatted %>%
  group_by(year) %>% tally()


### save to the `eora_cleaned` folder ----------------------------------------------------
fname <- paste0(dir.eora_cleaned, ind, '_matrix.xlsx'); fname
writexl::write_xlsx(x = df_m_formatted, path = fname)
```

```{r - flows_summary - net inflow}

flows <- df_m_formatted %>%
  select(-ctr) %>%
  filter(year == 2015) %>%
  select(-year) %>%
  column_to_rownames(var = 'iso3')

# Calculate total inflows (sum of columns)
inflows <- colSums(flows, na.rm = T)

# Calculate total outflows (sum of rows)
outflows <- rowSums(flows, na.rm = T)

# Calculate net inflows
net_inflows <- inflows - outflows

# Combine results into a data frame for clarity
flows_summary <- data.frame(
  Country = rownames(flows),
  Inflows = inflows,
  Outflows = outflows,
  NetInflows = net_inflows
)

```



## Flow map data 
  https://www.flowmap.blue/how-to-make-a-flow-map
  
```{r - flow data format}

df_l <- df_m_formatted %>%
  gather(4:ncol(.), key = 'dest', value = 'count') %>%
  rename('origin' = 'iso3', 
         'time' = 'year') %>%
  select(-ctr) %>%
  filter(time == 2015, 
         !origin %in% c('CIV', 'SUN'),
         !dest %in% c('CIV', 'SUN'),
         !is.na(count),
         count > 0) %>%
  select(origin, dest, count, time, everything())

f <- paste0(dir.dico, 'od_flow_', ind, '_2015.csv'); f
readr::write_csv(x = df_l, file = f)

```


