---
title: "Untitled"
author: "Yingjie"
date: "10/24/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---


# Set up

## packages
```{r}
### set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
setwd(dir)
setwd('../')
getwd()

library(tidyverse)
library(scales)
```


## help data
```{r}
## this data will be used to format flow data
load(file = paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')) # `temp_eora`
```



# Data

- HS92 (1995-2019, 2.18 Go)
- HS96 (1996-2019, 1.98 Go)
- HS02 (2002-2019, 1.62 Go)
- HS07 (2007-2019, 1.21 Go)
- HS12 (2012-2019, 799 Mo)
- HS17 (2017-2019, 300 Mo)


##  Metadata - BACI
### Country Code
```{r}
## BACI ############################################################################################
## - BACI provides disaggregated data on bilateral trade flows for more than 5000 products and 200 countries. 

# t	Year
# k	Product category (HS 6-digit code)
# i	Exporter (ISO 3-digit country code)
# j	Importer (ISO 3-digit country code)
# v	Value of the trade flow (in thousands current USD)
# q	Quantity (in metric tons)


dir.baci <- "D:/data/BACI/"

###
csv      <- paste0(dir.baci, 'country_codes_V202102.csv')
baci_ctr <- read_csv(csv) %>% as.data.frame()
```


### HS92
```{r}

### Trade data -------------------------------------------------------------------------------------
csv <- paste0(dir.baci, "BACI_HS92_V202102/", "BACI_HS92_Y2015_V202102.csv"); csv
# baci <- readr::read_csv(csv)
# head(baci)

csvs <- paste0(dir.baci, "BACI_HS92_V202102/", "BACI_HS92_Y", seq(2000, 2015, 5), "_V202102.csv"); csvs

# bacis <- data.frame()
# for (csv in csvs) {
#   print(csv)
#   baci  <- readr::read_csv(csv)
#   bacis <- rbind(bacis, baci)
# }
# save(bacis, file = paste0(dir.baci, 'BACI_HS92.RData'))


```


### HS07
```{r }
csvs <- paste0(dir.baci, "BACI_HS07_V202102/", "BACI_HS07_Y", c(2007, 2010, 2015), "_V202102.csv"); csvs

# bacis <- data.frame()
# for (csv in csvs) {
#   print(csv)
#   baci  <- readr::read_csv(csv)
#   bacis <- rbind(bacis, baci)
# }
# save(bacis, file = paste0(dir.baci, 'BACI_HS07_07_10_15.RData'))

```





## Pesticides

  To find the HS code for `Hazardous pesticides`
  Based on `RT_EN_README_Domain_Information_FAOSTAT item codes, HS codes and PIC.xlsx`, the HS codes are
    - 380850 (HS 07 – 12) Hazardous pesticides
    - 380859 (HS 17)      Hazardous pesticides
        
        
  **to do**
  - No 2000 and 2005? --> Yes, so far there is only data in HS07 (2007-2019)
  - Compare this data to the FAO data
```{r}

ind <- 'TradeMatrix_Pesticides_Hzd'

# f <- paste0(dir.baci, 'BACI_HS92.RData')          ## there is no data for 380850 and 380859
f <- paste0(dir.baci, 'BACI_HS07_07_10_15.RData')

load(file = f)

head(bacis)
ks <- bacis %>% distinct(k, .keep_all = F) %>%
  dplyr::filter(str_starts(k, '3808'))

pst_ij <- bacis %>% 
  # dplyr::filter(t == year) %>%
  ### - filter based on fishery code
  dplyr::filter(k %in% c('380850', '380859')) %>%
  ### - total flows among countries
  group_by(t, i, j) %>%
  dplyr::summarise_at(vars(v:q), sum, na.rm = TRUE) %>%
  ### - add country code 
  merge(., baci_ctr %>% dplyr::select(country_code, iso_3digit_alpha), by.x = 'i', by.y = 'country_code', all.x = T) %>%
  dplyr::rename(o = iso_3digit_alpha) %>%
  merge(., baci_ctr %>% dplyr::select(country_code, iso_3digit_alpha), by.x = 'j', by.y = 'country_code', all.x = T) %>%
  dplyr::rename(d = iso_3digit_alpha) %>%
  dplyr::select(-i, -j, -v) #%>%


### - To generate the same format as Eora in Matrix
df_formatted <- pst_ij %>%
  merge(x = temp_eora, y = ., by.x = c('year', 'iso3', 'to'), by.y = c('t', 'o', 'd'), all.x = T) %>%
  arrange(year, iso3, to) %>%
  dplyr::select(year, ctr, everything()) %>%
  spread(key = to, value = q) %>%
  arrange(year, iso3) %>%
  as.data.frame()

### - available date in each year --> 119 in 2015
pst_ij %>%
  spread(key = d, value = q) %>%
  group_by(t) %>% tally()

df_formatted %>%
  group_by(year) %>% tally()

### save to the `eora_cleaned` folder 
fname <- paste0('./Data/data_02_intermediate/dt02_flows/eora_cleaned/', ind, '.xlsx'); fname
writexl::write_xlsx(x = df_formatted, path = fname)
```





### Compare BACI to FAO trade data

  Based on the comparison, the two data sets align well with each other. 
  
```{r warning=FALSE}
pst_ij_export <- df_formatted %>%
  gather(key = 'to', value = 'value', 4:ncol(.)) %>%
  group_by(year, ctr, iso3) %>%
  dplyr::summarise(exp_baci = sum(value, na.rm = T))

dir.cleaned      <- './Data/data_02_intermediate/dt01_ctr_profile/xlsx/cleaned/'
xls <- paste0(dir.cleaned, 'Pesticides Trade_Export Quantity_4yrs.xlsx')
pst_fao_exp <- readxl::read_excel(xls) %>%
  gather(key = 'year', value = 'exp_fao', 3:ncol(.)) 



pst_compare <- merge(x = pst_fao_exp, y = pst_ij_export, by = c('year', 'ctr', 'iso3'), all.x = T) %>%
  dplyr::filter(year >= 2007)


pst_compare %>% ggplot(aes(x = exp_fao, y = exp_baci)) +
  geom_point(aes(color = year)) +
  # scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #               labels = trans_format("log10", math_format(10^.x))) +
  # scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #               labels = trans_format("log10", math_format(10^.x))) +
  geom_abline(linetype = "dashed") +
  theme_bw()

```

