---
title: "Spillover Analysis"
author: "Yingjie Li"
date: "2021-09-23"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


  This script aims to 
  
  1. Read the combined EORA indicator data (4 years data in one table), and calculate `net imports` for each indicator, each year, and each country. 
  
  2. Calculate `spillover score` and `spillover index`, according to SDSN's approach. 
  
  
# Set up

```{r}
### packages
source('./Code/_package list.R')
source('./Code/func_net_import_direction.R')

dir.eora         <- './Data/data_02_intermediate/dt02_flows/eora/'
dir.eora_cleaned <- './Data/data_02_intermediate/dt02_flows/eora_cleaned/'
dir.fig          <- './Data/Figure/'
```


#  Data 

## Neighbor matrix
```{r}
### load spatial relation: nearby or distant
ner_dst_matrix <- './Data/_neighbor_matrix/_input_ctr_matrix_neighbor.xlsx'
ner.matrix <- read_excel(path = ner_dst_matrix) %>% dplyr::select(-Country, -Neighbors) %>% column_to_rownames(var = "o")
diag(ner.matrix) <- 1
dst.matrix <- 1 - ner.matrix 
not.matrix <- (ner.matrix + dst.matrix)


### how many nearby and distant pairs?
ner_stat <- ner.matrix %>%
  rownames_to_column(var = 'o') %>%
  gather(key = 'd', value = 'scenario', 2:ncol(.)) %>%
  group_by(scenario) %>%
  tally()

ner_stat$n/sum(ner_stat$n) * 100 

### --> 96.9% distant pairs; 3.1% nearby pairs



### how many nearby or distant partners each country has? 
ner_stat_each <- ner.matrix %>%
  dplyr::mutate(n_nearby = rowSums(.),
                n_near_pct = n_nearby/189*100) %>% ## % are nearby
  dplyr::select(n_nearby, n_near_pct) %>%
  as.data.frame()

hist(ner_stat_each$n_nearby)
boxplot(ner_stat_each$n_nearby)
```




## Net imports and TNI Score by senarios

  Only indicators with decided *direction* were included in the calculation, while assign *NA* to those with un-decided direction. 
  - Need to update the `direction` table!
  
```{r }
### get the name list of all the spillovers and manually add `direction` info for further normalizing the values

xls.ls <- list.files(path = dir.eora_cleaned, pattern = 'xlsx$', full.names = T); xls.ls

spillover_name_list <- data.frame(ind = gsub('.xlsx', '', basename(xls.ls))) %>%
  # dplyr::mutate(shdb = str_detect(ind, pattern="SHDB")) %>%
  dplyr::mutate(direction = '') %>%  ## to determine if it is positive or negative impact
  as.data.frame()


### loop the normalization based on direction ------------------------------------------------------
xls <- paste0('./Data/data_02_intermediate/dt02_flows/', "spillover_name_list_directions.xlsx"); xls
direction <- readxl::read_excel(path = xls) %>%
  dplyr::filter(!is.na(ind) & !is.na(direction))
```





```{r - nearby only}

scenario <- 'fromNer'
matrix   <- ner.matrix


### put all spillover indicators in one table 
net_imports <- data.frame()


### this for loop can be used in other scenarios 
for (i in 1:nrow(direction)) {
  # print(i)
  ind <- direction[i, 1] %>% unlist() %>% paste(sep = '', collapse = ''); 
  drt <- direction[i, 2] %>% unlist() %>% paste(sep = '', collapse = '') %>% as.numeric()
  print(ind)
  print(drt)
  
  xls <- paste0(dir.eora_cleaned, ind, '.xlsx'); print(xls)
  df  <- readxl::read_excel(path = xls)  %>% 
    gather(key = 'd', value = 'value', 4:ncol(.)) %>%
    as.data.frame()
  
  df_wide <- df %>%
    dplyr::group_by(year) %>%
    arrange(year, iso3, d) %>%
    spread(key = 'd', value = 'value') %>%
    arrange(year, iso3) %>%
    as.data.frame()
    
  ### --> `df` under different scenarios
  mat <- matrix %>%
    rownames_to_column(var = 'o') %>%
    gather(key = 'd', value = 'scenario', 2:ncol(.))
  
  df_scenario <- merge(df, mat, by.x = c('iso3', 'd'), by.y = c('o', 'd'), all.x = T) %>%
    dplyr::mutate(value = value * scenario) %>%
    dplyr::select(-scenario) %>%
    spread(key = 'd', value = 'value') %>%
    arrange(year, iso3) %>%
    dplyr::select(year, ctr, iso3, everything()) %>%
    as.data.frame()
  
  net_in <- func_net_import_direction(df = df_scenario, direction = drt, frac = 0.025)
  net_imports <- rbind(net_imports, cbind(ind = ind, net_in))
  
}

net_imports$scenario <- scenario
net_imports_fromNer  <- net_imports
```

  
  
  
```{r - distant only}

scenario <- 'fromDst'
matrix   <- dst.matrix

### put all spillover indicators in one table 
net_imports <- data.frame()


### this for loop can be used in other scenarios 
for (i in 1:nrow(direction)) {
  # print(i)
  ind <- direction[i, 1] %>% unlist() %>% paste(sep = '', collapse = ''); 
  drt <- direction[i, 2] %>% unlist() %>% paste(sep = '', collapse = '') %>% as.numeric()
  print(ind)
  print(drt)
  
  xls <- paste0(dir.eora_cleaned, ind, '.xlsx'); print(xls)
  df  <- readxl::read_excel(path = xls)  %>% 
    gather(key = 'd', value = 'value', 4:ncol(.)) %>%
    as.data.frame()
  
  df_wide <- df %>%
    dplyr::group_by(year) %>%
    arrange(year, iso3, d) %>%
    spread(key = 'd', value = 'value') %>%
    arrange(year, iso3) %>%
    as.data.frame()
    
  ### --> `df` under different scenarios
  mat <- matrix %>%
    rownames_to_column(var = 'o') %>%
    gather(key = 'd', value = 'scenario', 2:ncol(.))
  
  df_scenario <- merge(df, mat, by.x = c('iso3', 'd'), by.y = c('o', 'd'), all.x = T) %>%
    dplyr::mutate(value = value * scenario) %>%
    dplyr::select(-scenario) %>%
    spread(key = 'd', value = 'value') %>%
    arrange(year, iso3) %>%
    dplyr::select(year, ctr, iso3, everything()) %>%
    as.data.frame()
  
  net_in <- func_net_import_direction(df = df_scenario, direction = drt, frac = 0.025)
  net_imports <- rbind(net_imports, cbind(ind = ind, net_in))
  
}



net_imports$scenario <- scenario
net_imports_fromDst  <- net_imports
``` 




```{r - Warning !!!}
### In the scenario of only nearby interaction, we need to consider the 'burden' from `distant` interactions (Be careful here!) 

```





#### - Save R data
```{r}

net_imports_fromDst_fromNer <- rbind(net_imports_fromNer, net_imports_fromDst)


## shdb do not have data in 2000, 2005, 2010
net_imports_bind_check <- net_imports_fromDst_fromNer %>%
  dplyr::filter(!year %in% c(2000, 2005, 2010, 2015)) %>%
  dplyr::distinct(ind, year, .keep_all = T)

### save the data
n_ft_ind <- length(unique(net_imports_fromDst_fromNer$ind));  n_ft_ind
fname <- paste0(dir.eora_cleaned, 'net_imports_fromDst_fromNer_', n_ft_ind, '.RData'); fname
save(net_imports_fromDst_fromNer, file = fname)
```






