---
title: "Untitled"
author: "Yingjie"
date: "9/23/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


This script aims to 

  1. do the calculation based on the approach used in Xu and Li et al NS paper, and 
  
  2. compare and validate with results based on new loop-approach.  


# Setup
```{r}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
# remove(list = ls())


dir.eora_cleaned <- './Data/data_02_intermediate/dt02_flows/eora_cleaned/'
dir.fig     <- './Data/Figure/'

library(dplyr)
library(ggplot2)
library(cowplot)


### This code was used in Xu & Li et al 2020 -----------------------------
source('./Code/Func_Net_Import_Cal.R')
```



```{r - SI data}
load('./Data/Ancillary_Data_ISO3code_shp.RData') ## iso_eora, pop, gdp, shp, grp

grp_id <- grp %>%
  dplyr::mutate(group_id = as.numeric(gsub("\\D", "", inc_group))) %>%
  dplyr::select(iso_a3, group_id) %>%
  dplyr::filter(!is.na(iso_a3)) %>%
  as.data.frame()
```



# Data

## neighbor matrix

```{r}

# load spatial relation: nearby or distant
ner_dst_matrix <- './Data/_neighbor_matrix/_input_ctr_matrix_neighbor.xlsx'
matrix.ner <- read_excel(path = ner_dst_matrix) %>% dplyr::select(-Country, -Neighbors) %>% column_to_rownames(var = "o")
diag(matrix.ner) <- 1
matrix.dst <- 1 - matrix.ner 
matrix.not <- (matrix.ner + matrix.dst)
```


## Read in from excel
```{r}

# df.xlsx <- "0_capital.xlsx"
df.xlsx <- paste0(dir.eora_cleaned, "Water.xlsx");          SDG <- 'SDG6'
# df.xlsx <- "7_ene.xlsx";          SDG <- 'SDG7'
# df.xlsx <- "8_mat.xlsx";          SDG <- 'SDG8'
# df.xlsx <- "9_co2.xlsx";          SDG <- 'SDG9'
# df.xlsx <- "12_mat-perCap.xlsx";  SDG <- 'SDG12'
# df.xlsx <- "13_co2-forest.xlsx";  SDG <- 'SDG13'
# df.xlsx <- "15_for.xlsx";         SDG <- 'SDG15'




#### cal net import under different scenarios: (1) no trade, (2) only nearby trade, (3) only distant trade
net_in_not <- Net_Import_Cal(df.xlsx = df.xlsx, matrix = matrix.not)
net_in_ner <- Net_Import_Cal(df.xlsx = df.xlsx, matrix = matrix.ner)
net_in_dst <- Net_Import_Cal(df.xlsx = df.xlsx, matrix = matrix.dst)

```


```{r}

### compare to the New calculation approach -----
net_in_ner_new <- net_imports_fromNer %>% 
  dplyr::filter(ind == 'Water') %>%
  dplyr::select(year, iso3, net_in) %>%
  arrange(year, iso3) %>%
  spread(key = year, value = net_in) %>%
  column_to_rownames(var = 'iso3')

dif <- net_in_ner - net_in_ner_new
```







## Indicator calculation
### 6.4.1. WU/GDP
```{r}
### GDP data (10^6 $)


GDP_c     <- gdp_rel_not %>% 
  dplyr::filter(!iso3 %in% c("ANT", "ROW")) %>%
  dplyr::rename(rel_c = gdp_rel, not_c = gdp_not) %>%
  gather(key = scenario, value = value, 3:4) %>% spread(key = 'year', value = 'value')
GDP_c_rel <- GDP_c %>% filter(scenario == 'rel_c') %>% dplyr::select(-scenario) %>% column_to_rownames(var = "iso3")
GDP_c_not <- GDP_c %>% filter(scenario == 'not_c') %>% dplyr::select(-scenario) %>% column_to_rownames(var = "iso3")
# GDP_c_ner <- GDP_c %>% filter(scenario == 'ner_c') %>% dplyr::select(-1, -17)
# GDP_c_dst <- GDP_c %>% filter(scenario == 'dst_c') %>% dplyr::select(-1, -17)


### WU - water use (1000 m3 = 10^3 m3 ???)
loc <- rel %>% 
  spread(key = 'year', value = 'value') %>%
  dplyr::filter(!iso3 %in% c("ANT", "ROW")) %>%
  column_to_rownames(var = "iso3")
loc <- loc[ order(row.names(loc)), ] %>% as.data.frame()


### cal GDP/WU ($/m3), to keep the row names, we cal using WU/GDP
df_rel_c <- GDP_c_rel / (loc + 0     )
df_not_c <- GDP_c_not / (loc + net_in_not)
df_ner_c <- GDP_c_ner / (loc + net_in_dst) ## if only nearby trade, then add burden from distant-trade
df_dst_c <- GDP_c_dst / (loc + net_in_ner)

## normalization using the function - Func_norm_sdg_score
score_c1 <- Func_norm_sdg_score_bad(df_rel_c, df_not_c, 
                                    #df_ner_c, df_dst_c,
                                    bottom = 0.025, top = 0.975)
## look at the data distribution
score_c1_long <- score_c1 %>% gather(year, value, 2:16)

ggplot(score_c1_long, aes(x = value)) +
  geom_histogram() +
  facet_wrap(~scenario) +
  theme_bw() +
  ggtitle('Indicator 6.4.1')



### save the data for side use ---------------------------------------------------
getwd()
normalized_score <- score_c1
fname <- paste0(dir.root, '/data/data_myRData/normalized_score__', 'gdp_per_water', '.RData'); fname
save(normalized_score, file = fname)
```



### 6.4.2. WU/WR
```{r}
### WU - water use (1000 m3 = 10^3 m3 ???)
loc <- read_excel(df.xlsx, sheet = 'S7.net_import_nearby', col_names = T, range = ("R1:AI42")) 

loc <- as.data.frame(loc[ , names(loc) != "...17"]) 
# first col as row name
loc <- data.frame(loc[,-1], row.names=loc[,1])
loc <- loc[ order(row.names(loc)), ] %>% as.data.frame()


### total water resources (10^9 m3)
local_total <- (loc[, 16]*10^9) 

df_rel_c <- (loc[, -16] + 0     ) *10^3/local_total
df_not_c <- (loc[, -16] + net_in_not) *10^3/local_total
df_ner_c <- (loc[, -16] + net_in_dst) *10^3/local_total 
# ************** (loc + net distant imports) means SDG under only  nearby trade scenario
df_dst_c <- (loc[, -16] + net_in_ner) *10^3/local_total 
# ************** (loc + net nearby  imports) means SDG under only distant trade scenario


## normalization using the function - Func_norm_sdg_score
## load the function
source(paste0(dir, "/Func_norm_sdg_score_good_bad.R"))

score_c2 <- Func_norm_sdg_score_bad(df_rel_c, df_not_c, df_ner_c, df_dst_c,
                                    bottom = 0.025, top = 0.975)
## look at the data distribution
score_c2_long <- score_c2 %>% gather(year, value, 2:16)
ggplot(score_c2_long, aes(x = value)) +
  geom_histogram() +
  facet_wrap(~scenario) +
  theme_bw() +
  ggtitle('Indicator 6.4.2')


### save the data for side use ---------------------------------------------------
getwd()
normalized_score <- score_c2
fname <- paste0(dir.root, '/data/data_myRData/normalized_score__', 'water_stress', '.RData'); fname
save(normalized_score, file = fname)

```




### Comapre two indicators

```{r}

score_c1_long_rel <- score_c1 %>% filter(scenario=='rel_c') %>%
  dplyr::mutate(MeanScore = rowMeans(select(., `1995`:`2009`))) %>%
  # arrange(rev(MeanScore)) %>%
  as.data.frame() %>%
  dplyr::mutate(colorID = rank(MeanScore)) %>%
  gather(year, value, 2:16) %>%
  dplyr::mutate(id = row_number())

score_c2_long_rel <- score_c2 %>% filter(scenario=='rel_c') %>%
  dplyr::mutate(MeanScore = rowMeans(select(., `1995`:`2009`))) %>%
  dplyr::mutate(colorID = rank(MeanScore)) %>%
  gather(year, value, 2:16)%>%
  dplyr::mutate(id = row_number())
str(score_c1_long_rel)


ggplot(score_c2_long_rel, aes(x = reorder(nation, colorID), y = value, color =as.factor(colorID))) + 
  # geom_point() + 
  geom_boxplot() +
  theme_bw() +
  # ggtitle('Indicator 15.2.1') +
  coord_flip() +
  theme(legend.position="none")

### plot these 2 indicators in one
score_c12_long_rel <- merge(x = score_c1_long_rel, y = score_c2_long_rel, by = 'id') %>%
  gather(indicator, value, c(value.x, value.y))


unique(score_c12_long_rel$indicator)
label_names <- c(`value.x` = "Indicator 6.4.1", `value.y` = "Indicator 6.4.2")

ggplot(score_c12_long_rel, 
       aes(x = reorder(nation.x, colorID.x), 
           y = value, 
           color =as.factor(colorID.x))) + 
  # geom_point() + 
  geom_boxplot() +
  theme_bw() +
  # ggtitle('SDG 15') +
  ylab('SDG score') +
  xlab('Country') +
  coord_flip() +
  facet_wrap(~indicator, scales = 'free_x', labeller = as_labeller(label_names))+
  theme(legend.position="none")

getwd()
fname <- paste0(dir.figures, '/compare_2_indicators_', SDG, '-2020.png'); fname
ggsave(filename = fname,  plot = last_plot(), width = 12, height = 6, units = "in", dpi = 500)

```




## Mean SDG score
Take the mean of these two indicators
```{r}
SDG_score_c <- data.frame(score_c1[,1], 
                          (score_c1[,2:16] + score_c2[,2:16])/2, 
                          score_c1[,17:18]) %>% as.data.frame()
names(SDG_score_c) <- c('nation', seq(1995, 2009), "scenario", "group")



# save as csv for cal SDGc
out.dir <- './_output_score'
fname <- paste0(out.dir, '/score_c_', gsub('.xlsx', '', df.xlsx), '.csv'); fname
write.csv(x = SDG_score_c, file = fname, row.names = T)

```



## Diff between dst and ner; by global/ group

```{r}

source(paste0(dir, "/single_diff_func_c_approach.R"))
source(paste0(dir, "/single_diff_func_b_approach.R"))
 
single_diff_func_c(SDG_score_c)
# single_diff_func_b(score_b)

```





# Data_Descriptive_Stats

## data
```{r}
library(tidyverse)
library(reshape2) # expand long table to wide one
library(scales) # to access break formatting functions


s1 <- read_excel(df.xlsx, sheet = 1, col_names = T, range = "B1:AP616")

names(s1)
s3 = s1

# add country name as a new col
countries <- colnames(s1[1:41])
colnames(s3) <- countries; length(countries)

# add year seq to data frame 
year_seq = rep(seq(1995, 2009, by=1), each = 41); year_seq
s3$year <- year_seq

### total imports
s4 <- s3 %>% group_by(year) %>% summarise_all(funs(sum))
s5 <- as.data.frame(t(s4[,-1])) ## tanspose data frame and keep the first col as new header
colnames(s5) <- seq(1995, 2009, by=1)
s5 <- s5[ order(row.names(s5)), ]

### total  exports
s6 <- s3 %>%
  mutate(row.sum   = rowSums(s3[,1:41]),
         year      = year_seq,
         countries = rep(countries, times = 15)) %>% # times=15, each = 15
  dcast(countries ~ year, value.var = 'row.sum')    # year as new col names
s6 <- data.frame(s6[,-1], row.names=s6[,1])
s6.1 <- s6[ order(row.names(s6)), ]


### input data for plot
ip <- s4[-1] %>% gather(key = 'nation', value = 'value') ## wide to long
ep <- as.data.frame(t(s6)) %>% gather(key = 'nation', value = 'value')

str(ip)

ip.mean <- ip %>% group_by(nation) %>% dplyr::summarise(mean = mean(value))
ip.new  <- merge(x = ip, y = ip.mean, by = 'nation', all.x = T)

ip_ep <- rbind(
  cbind(trade = 'imports', ip),
  cbind(trade = 'exports', ep)
)
ip_ep.new  <- merge(x = ip_ep, y = ip.mean, by = 'nation', all.x = T)
```


## plot

```{r}
lab <- bquote('Virtual water transfered ('~1000~m^3*')')

ggplot(data = ip_ep.new, aes(x = reorder(nation, mean), y = value, color = trade)) + ## log10(value)
  # geom_point() + 
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  theme_bw() +
  # ggtitle('Indicator 15.2.1') +
  # coord_flip() +
  # theme(legend.position="none") +
  # scale_color_hue(direction = -1) +
  # scale_colour_brewer(direction = -1)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  theme(legend.position = c(0.8, 0.2)) +
  ylab(lab) +
  xlab('Country')

fname <- paste0(dir.figures, '/_Data_Descriptive_Stats_', SDG, '-2020.png'); fname
ggsave(filename = fname, plot = last_plot(), width = 7, height = 5.5, units = "in", dpi = 500)

```







