---
title: "Untitled"
author: "Yingjie"
date: "6/30/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Overview

  To get the output for the analysis, we only need to run these two chunks
  - `#### --> All indicators`
  - `# 4. Plot network for each TNI`
  
  Other sections/chunks are for testing and selecting the best viz option. 
  


## 11. Map network

Plot a network graph of nodes with geographic coordinates on a map.
Author: Markus Konrad <markus.konrad@wzb.eu>
May 2018
https://gist.github.com/internaut/a9a274c72181eaa7f5c3ab3a5f54b996#file-networkmap-r

This script shows three ways of plotting a network graph on a map.
The following information should be visualized (with the respective aestethics added):

  * graph nodes with:
    * position on map -> x,y position of the node
    * node weight (degree of the node) -> node size
    * node label -> also x,y position of the node
  * edges between nodes with:
    * edge weight -> edge width
    * edge category -> edge color
  
```{r}

### To clear your environment 
remove(list = ls())

### set work dir
# path <- rstudioapi::getSourceEditorContext()$path
# dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
# setwd(dir)
# setwd('..') # set directory by one folder up
getwd()

library(assertthat)
library(dplyr)
library(purrr)
library(igraph)
library(ggplot2)
library(ggraph)
library(ggmap)
```


```{r - data example}
# -------------------------------------- #
# Preparation: generate some random data #
# -------------------------------------- #

country_coords_txt <- "
   1     3.00000  28.00000       Algeria
   2    54.00000  24.00000           UAE
   3   139.75309  35.68536         Japan
   4    45.00000  25.00000 'Saudi Arabia'
   5    9.00000   34.00000       Tunisia
   6     5.75000  52.50000   Netherlands
   7   103.80000   1.36667     Singapore
   8   124.10000  -8.36667         Korea
   9    -2.69531  54.75844            UK
  10    34.91155  39.05901        Turkey
  11  -113.64258  60.10867        Canada
  12    77.00000  20.00000         India
  13    25.00000  46.00000       Romania
  14   135.00000 -25.00000     Australia
  15    10.00000  62.00000        Norway"

# nodes come from the above table and contain geo-coordinates for some
# randomly picked countries
nodes <- read.delim(text = country_coords_txt, header = FALSE,
                    quote = "'", sep = "",
                    col.names = c('id', 'lon', 'lat', 'name')) %>%
  dplyr::rename('label' = 'name')

# edges: create random connections between countries (nodes)
set.seed(123)

N_EDGES_PER_NODE_MIN <- 1
N_EDGES_PER_NODE_MAX <- 4
N_CATEGORIES <- 4

edges <- map_dfr(nodes$id, function(id) {
  n <- floor(runif(1, N_EDGES_PER_NODE_MIN, N_EDGES_PER_NODE_MAX+1))
  to <- sample(1:max(nodes$id), n, replace = FALSE)
  to <- to[to != id]
  categories <- sample(1:N_CATEGORIES, length(to), replace = TRUE)
  weights <- runif(length(to))
  data_frame(from = id, to = to, weight = weights, category = categories)
})

edges <- edges %>% mutate(category = as.factor(category))


nodes.example <- nodes
edges.example <- edges
```


**...Tips...** 
edges
  - weight to be normalized to 0-1
  

```{r - your data - simplify}
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'network_data_for_viz.RData')
load(f) ## `links_scaled_all_Net`, `grp_update_net`, `pop`

### load capital city data
f <- "./Data/capital_city_data.rds"
capital_city_data <- readRDS(file = f)

### choose one of the indicators to compare 
ind <- 'Landuse_forest'
ind <- 'total'
  
source('https://raw.githubusercontent.com/Yingjie4Science/functionX/main/scaling_normalization_Pro.R')

edges <- links_scaled_all_Net %>%
  dplyr::select(all_of(c('from', 'to', ind))) %>%
  ##' need to scale the range of weight to 0-1
  dplyr::mutate(weight = func_norm_minmax(x=total, a=0, b=4), 
                weight = ifelse(weight<0, 0, weight)) %>%
  ##' look at the top 15 flows
  arrange(desc(weight)) %>%
  head(105) %>% ## 55
  ## remove nodes with too few links (< 5 links)
  group_by(from) %>%
  dplyr::mutate(freq_from = n()) %>%
  dplyr::filter(freq_from >= 1) %>%  
  dplyr::select(-freq_from) %>%
  ungroup() %>%
  
  dplyr::mutate(category = 1) %>%
  as.data.frame()


nodes <-
  data.frame(iso3 = c(edges$from, edges$to)) %>%
  dplyr::mutate(iso3 = as.character(iso3)) %>%
  dplyr::distinct(iso3, .keep_all = T) %>%

  merge(x = ., 
        y = grp_update_net, 
        by.x = 'iso3', by.y = 'iso_a3', all.x = T) %>%
  dplyr::distinct(iso3, .keep_all = T) %>%
  dplyr::mutate(group = as.factor(group_income2)) %>%
  merge(x = ., 
        y = pop %>% dplyr::select(iso3, `2015`), 
        by.x = 'iso3', by.y = 'iso3', all.x = T) %>%
  arrange(iso3) %>%
  dplyr::mutate(node_type  = ifelse(group == 'High income', 1, 0), 
                type.label = group) %>%
  dplyr::rename('id' = 'iso3', 'size_node' = '2015', 'label' = 'name') %>%
  dplyr::mutate(label = id) %>%
  
  ### add capital city's lon and lat info
  left_join(., capital_city_data, by = c('id' = 'iso3')) %>%
  dplyr::rename('lon' = 'long') %>%
  as.data.frame()
```



```{r}
# create the igraph graph object

g <- graph_from_data_frame(edges, directed = F, vertices = nodes)

# --------------------------------------------------------------------- #
# Common data structures and ggplot objects for all the following plots #
# --------------------------------------------------------------------- #

# create a data frame for plotting the edges
# join with nodes to get start and end positions for each
# edge (x, y and xend, yend)

edges_for_plot <- edges %>%
  inner_join(nodes %>% select(id, lon, lat), by = c('from' = 'id')) %>%
  rename(x = lon, y = lat) %>%
  inner_join(nodes %>% select(id, lon, lat), by = c('to' = 'id')) %>%
  rename(xend = lon, yend = lat)

assert_that(nrow(edges_for_plot) == nrow(edges))

# use the node degree for scaling the node sizes
nodes$weight = igraph::degree(g)

# common plot theme
maptheme <- theme(
  panel.border = element_blank(),
  panel.grid = element_blank()) + 
  theme(axis.text = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.title = element_blank()) +
  theme(legend.position = "bottom") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_rect(fill = "transparent")) +
  theme(plot.margin = unit(c(0, 0, 0.5, 0), 'cm'))

# common polygon geom for plotting the country shapes
country_shapes <- geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group),
                               fill = "transparent", color = "gray50", size = 0.01, alpha = .8)
# common coordinate system for all the following plots
mapcoords <- coord_fixed(xlim = c(-150, 180), ylim = c(-55, 80))


color_label <- "black"
color_label <- "transparent"
```



### Solution 1: ggplot + ggmap only
```{r}
# ------------------------------- #
# Solution 1: ggplot + ggmap only # -------------
# ------------------------------- #

# try to plot with scaled edge widths and node sizes
# this will fail because we can only use the "size" aesthetic twice
ggplot(nodes) + 
  country_shapes +
  geom_curve(aes(x = x, y = y, xend = xend, yend = yend,     # draw edges as arcs
                 color = category, size = weight/5),
             arrow = arrow(length = unit(0.2, "in"), ends = 'last', 
                           angle = '20', # default = 30
                           type = 'open'),
             data = edges_for_plot, curvature = 0.33, alpha = 0.5) +
  scale_size_continuous(guide = FALSE, range = c(0.25, 2)) + # scale for edge widths
  geom_point(aes(x = lon, y = lat, size = weight),           # draw nodes
             shape = 21,   
             fill = 'white', color = 'black', stroke = 0.5) +
  scale_size_continuous(guide = FALSE, range = c(1, 6)) +    # scale for node size
  geom_text(aes(x = lon, y = lat, label = label),             # draw text labels
            hjust = 0, nudge_x = 1, nudge_y = 4,
            size = 3, color = color_label) +
  mapcoords + maptheme



##' Results in warning: "Scale for 'size' is already present. 
##'   Adding another scale for 'size', which will replace the existing scale."
##' now a plot with static node size:

ggplot(nodes) + 
  country_shapes +
  geom_curve(aes(x = x, y = y, xend = xend, yend = yend,     # draw edges as arcs
                 color = category, size = weight),
             # arrow = arrow(length = unit(0.1, "in"), ends = 'last', 
             #               angle = '20', # default = 30
             #               type = 'open'),
             data = edges_for_plot, curvature = 0.33, alpha = 0.5) +
  scale_size_continuous(guide = FALSE, range = c(0.25, 2)) + # scale for edge widths
  geom_point(aes(x = lon, y = lat),                          # draw nodes
             shape = 21, size = 3,
             fill = 'white', color = 'black', stroke = 0.5) +
  geom_text(aes(x = lon, y = lat, label = label),             # draw text labels
            hjust = 0, nudge_x = 1, nudge_y = 4,
            size = 3, color = color_label) + # , fontface = "bold"
  mapcoords + maptheme
```


### Solution 2: ggplot2 + ggmap + ggraph
```{r}
# ------------------------------------ #
# Solution 2: ggplot2 + ggmap + ggraph #--------------
# ------------------------------------ #

# prepare layout: use "manual" layout with geo-coordinates
node_pos <- nodes %>% select(lon, lat) %>% rename(x = lon, y = lat)
lay <- create_layout(g, 'manual', node.positions = node_pos)
assert_that(nrow(lay) == nrow(nodes))

# use the node degree for scaling the node sizes
lay$weight <- igraph::degree(g)

ggraph(lay) + country_shapes +
  geom_edge_arc(aes(color = category, edge_width = weight,   # draw edges as arcs
                    circular = FALSE),
                data = edges_for_plot, curvature = 0.33, alpha = 0.5) +
  scale_edge_width_continuous(range = c(0.5, 2),             # scale for edge widths
                              guide = FALSE) +
  geom_node_point(aes(size = weight), shape = 21,            # draw nodes
                  fill = "white", color = "black",
                  stroke = 0.5) +
  scale_size_continuous(range = c(1, 6), guide = FALSE) +    # scale for node widths
  geom_node_text(aes(label = name), repel = TRUE, size = 3,
                 color = color_label, fontface = "bold") +
  mapcoords + maptheme
```



### Solution 3: the hacky way (overlay several ggplot "plot grobs")
```{r}
# --------------------------------------------------------------- #
# Solution 3: the hacky way (overlay several ggplot "plot grobs") # -----------
# --------------------------------------------------------------- #

theme_transp_overlay <- theme(
  panel.background = element_rect(fill = "transparent", color = NA),
  plot.background = element_rect(fill = "transparent", color = NA), 
  panel.border = element_blank(),
)

# the base plot showing only the world map
p_base <- ggplot() + country_shapes + mapcoords + maptheme

# first overlay: edges as arcs
p_edges <- ggplot(edges_for_plot) +
  geom_curve(aes(x = x, y = y, xend = xend, yend = yend,     # draw edges as arcs
                 color = category, size = weight),
             curvature = 0.33, alpha = 0.5) +
  scale_size_continuous(guide = FALSE, range = c(0.5, 2)) +  # scale for edge widths
  mapcoords + maptheme + theme_transp_overlay +
  theme(legend.position = c(0.5, -0.1), legend.direction = "horizontal")

# second overlay: nodes as points
p_nodes <- ggplot(nodes) +
  geom_point(aes(x = lon, y = lat, size = weight),
             shape = 21, fill = "white", color = "black",    # draw nodes
             stroke = 0.5) +
  scale_size_continuous(guide = FALSE, range = c(1, 6)) +    # scale for node size
  geom_text(aes(x = lon, y = lat, label = label),             # draw text labels
            hjust = 0, nudge_x = 1, nudge_y = 4,
            size = 3, color = color_label) +
  mapcoords + maptheme + theme_transp_overlay

# combine the overlays to a full plot

# proper positioning of the grobs can be tedious... I found that
# using `ymin` works quite well but manual tweeking of the
# parameter seems necessary

p <- p_base +
  annotation_custom(ggplotGrob(p_edges), ymin = -74) +
  annotation_custom(ggplotGrob(p_nodes), ymin = -74)

print(p)

```

