---
title: "Untitled"
author: "Yingjie Li"
date: "6/4/2020"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dir and packages
```{r}
# set work dir ------------------------------------------------------------

### set work dir
path <- rstudioapi::getSourceEditorContext()$path; path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir

setwd(dir)
getwd()

library(readxl)
library(writexl)
library(tidyverse)
library(reshape2)
library(stringr)
# library(here)
library(Rmisc)
library(scales)
library(sf)
library(ggplot2)
library(cowplot)
library(tmap)
library(ggpubr)
```


# Raw data
## Data cleaning
```{r}
### read data ---------------------------------------------------------------
# df <- read.csv('./EDUN/EDUN_DATA_NATIONAL.csv', stringsAsFactors = F)
# save(df, file = './EDUN/EDUN_DATA_NATIONAL.RData')
load(file = './EDUN/EDUN_DATA_NATIONAL.RData')
names(df)
## "INDICATOR_ID": origin of the student flows
## "COUNTRY_ID":   destination of the student flows
## "YEAR"         
## "VALUE"

### load the indicator code and description
lables <- read.csv('./EDUN/EDUN_LABEL.csv', stringsAsFactors = F) 


### Data cleaning -----------------------------------------------------------
### the country name list in the raw data
ctr_raw <- read.csv('./EDUN/EDUN_COUNTRY.csv', stringsAsFactors = F) 
names(ctr_raw) <- c("COUNTRY_ID", "COUNTRY_NAME_EN")
### some ctr names are missing, we added these from other sources
ctr_raw_add <- readxl::read_excel('./EDUN/__additional_iso3_code_EDUN_COUNTRY.xlsx')
### append these data together
ctr_raw <- rbind(ctr_raw, ctr_raw_add)

### unify the ctr names
ctr_raw <- ctr_raw %>%
  mutate(COUNTRY_NAME_EN = gsub('Cabo Verde', 'Cape Verde', COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('Eswatini', 'Swaziland',    COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('United States of America', 
                                'United States', COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('China\\, Hong Kong Special Administrative Region', 
                                'Hong Kong\\, Special Administrative Region of China', 
                                COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('China\\, Macao Special Administrative Region', 
                                'Macao\\, Special Administrative Region of China', 
                                COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('Iran \\(Islamic Republic of\\)', 
                                'Islamic Republic of Iran', COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('Czechia', 'Czech Republic', COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('United Kingdom of Great Britain and Northern Ireland', 
                                'United Kingdom', COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('Micronesia \\(Federated States of\\)', 
                                'Federated States of Micronesia', COUNTRY_NAME_EN),
         COUNTRY_NAME_EN = gsub('Venezuela \\(Bolivarian Republic of\\)', 
                                'Bolivarian Republic of Venezuela', COUNTRY_NAME_EN)
  )
```




## Sending sys
```{r}
### for sending sys ------------------------------------------------------------
names(df)
## "INDICATOR_ID": origin of the student flows

ctr_send <- lables %>%
  ### ctr_send info is in the "INDICATOR_LABEL_EN" column 
  dplyr::filter(stringr::str_detect(INDICATOR_LABEL_EN, 'Students from')) %>%
  ### remove unnecessary info, to extract ctr names
  dplyr::mutate(
    origin = gsub('South America|Asia|Europe|Oceania|Caribbean and Central America|North America|Africa\\:', 
                  '', INDICATOR_LABEL_EN),
    origin = gsub('\\, both sexes \\(number\\)|Students from the |Students from ', 
                  '', origin),
    origin = gsub('\\:', '', origin),
    origin = gsub('former Yugoslav Republic of Macedonia', 'North Macedonia', origin),
    origin = trimws(origin)) %>%
  ### join the iso code by country full names
  left_join(x = ., y = ctr_raw, by = c("origin" = "COUNTRY_NAME_EN")) %>%
  ### any NA in the ctr iso code?
  arrange(!is.na(COUNTRY_ID)) %>%
  ### fix this iso3 code
  dplyr::mutate(
    COUNTRY_ID = ifelse(INDICATOR_ID==26433, 'CIV', COUNTRY_ID),
    ### 'unknown countries' as 'rest of world'
    COUNTRY_ID = ifelse(origin=='unknown countries', 'ROW', COUNTRY_ID) 
    ) %>%
  ### any NA in the ctr iso code?
  arrange(!is.na(COUNTRY_ID), COUNTRY_ID) 

### rename the ctr_send
names(ctr_send)
names(ctr_send) <- c("INDICATOR_ID", "INDICATOR_LABEL_EN", "origin", "send_iso")
# unique(ctr_send$INDICATOR_LABEL_EN)
# unique(ctr_send$origin)
# head(ctr_send, 3)
# str(ctr_send)

### how many sending sys?
n_sending <- length(unique(ctr_send$origin)); n_sending
### the sending sys list
ctr_send_ls <- ctr_send$INDICATOR_ID
```


## Receiving sys
```{r}
### for receiving sys -----------------------------------------------
head(df,3)

dt <- df %>%
  ### only keep the first 3 columns that are useful
  dplyr::select(INDICATOR_ID:VALUE) %>%
  # filter(YEAR >= 2013) %>%
  ### rename the col for receiving sys
  dplyr::rename(receive_iso = COUNTRY_ID) %>%
  # filter(INDICATOR_ID %in% c(26528)) %>% ##
  ## filter out the internationally mobile student data, based on INDICATOR_ID, which contains the names of sending ctr
  dplyr::filter(INDICATOR_ID %in% ctr_send_ls) %>% 
  spread(key = YEAR, value = VALUE) 

### get the unique receiving ctr name list
ctr_receive <- dt %>% 
  distinct(receive_iso) %>% as.data.frame() %>% 
  arrange(!is.na(receive_iso), receive_iso)

### how many receiving ctr?
n_receiving <- length(unique(ctr_receive$receive_iso)); n_receiving
print(paste0('There are ', n_receiving, ' receiving countries.'))
```



## Double-check on data
  The above data is downloaded as a bulk zip file, including many indicators. Here we go to the same webpage and download a subset of the data in another way, in which we can choose certain country and indicator. 
  We are going to use this small subset to validate and double-check on the bulk data we clearned. 
  Here, we downloaded the data for "China (chn) --> Other 100+ rest of world countries (row)". 
```{r}
### subset data, which was manually downloaded, for validation
chn_to_row <- readxl::read_excel(
  path = './EDUN/__manual_download_China2otherCountries.xlsx', skip = 3) %>%
  slice(-c(1, 243:288)) %>%
  select(-2) %>%
  dplyr::rename(country_name = Time) %>%
  left_join(x = ., y = ctr_raw, by = c('country_name' = 'COUNTRY_NAME_EN')) %>%
  select(COUNTRY_ID, country_name, 2:8) %>%
  arrange(COUNTRY_ID)  ## these are the receiving countries

# writexl::write_xlsx(chn_2_all, './EDUN/__country_names.xlsx')
print(paste0('There are ', length(unique(chn_to_row$COUNTRY_ID)),' receiving countries in manually download data subset'))

### compare to the receiving ctr list from the data 
print(paste0('There are ', n_receiving,   ' sending countries.'))
print(paste0('There are ', length(unique(chn_to_row$COUNTRY_ID)) - n_receiving, ' sending countries are missing in the bulk dataset, compared to manually downloaded data.'))


### to check which receiving ctr is missing ------------------------------------
ctr_receive_check <- ctr_receive %>%
  dplyr::mutate(check = 1) %>% ## 1 means ctr names in the data
  ### right join with the manual-download data, if check is NA, then these ctr are missing
  right_join(chn_to_row, by = c('receive_iso' = 'COUNTRY_ID')) %>%
  ### place NA on the top 
  arrange(!is.na(check), desc(check), receive_iso) ## desc
#### it seems some countries did NOT receive any students, and this is why these ctr are missing in the bulk data

### to check if sending and receiving ctr are the same list --------------------
check_sending_receiving <- ctr_receive %>%
  dplyr::mutate(check = 1,
                ### make a copy here
                receive_iso3 = receive_iso) %>%
  merge(x = ctr_send[, 3:4], y = ., 
        by.x = 'send_iso', by.y = 'receive_iso', all = T) %>%
  arrange(!is.na(check), desc(check), send_iso) ## desc
```



# Matrix as the output
  All looks good, and we can change the data format to a matrix-like data frame. 
```{r}
dt_mx_long <- ctr_send %>%
  right_join(x = ., y = dt, by = 'INDICATOR_ID') %>%
  ### there are some repeated send-receive ctr pairs, need to sum up 
  group_by(send_iso, receive_iso) %>%
  dplyr::summarise_at(.vars = as.character(seq(1998, 2019)),
               sum, na.rm = TRUE) %>%
  as.data.frame() %>%
  arrange(send_iso, receive_iso)
str(dt_mx_long)

### change to a send-receive matrix-like data frame ----------------------------
dt_mx_wide <- dt_mx_long %>%
  ### remove the first 3 cols, which are not useful here
  # dplyr::select(-c(1:3)) %>%
  ### gather years to one col
  gather(key = 'year', 'num', `1998`:`2019`) %>%
  as.data.frame() %>%
  ### spread receiving ctr 
  pivot_wider(names_from = receive_iso, values_from = num) %>%
  as.data.frame() %>%
  arrange(year, send_iso)
str(dt_mx_wide)


### take a look at the data for one year ---------------------------------------
dt_mx_2015 <- dt_mx_wide %>%
  # dplyr::select(send_iso, receive_iso, `2015`) %>%
  dplyr::filter(year == 2015) %>%
  arrange(send_iso)

### save data to local
writexl::write_xlsx(x = dt_mx_wide, path = './output_student_flows_1998_2019_UN_wide.xlsx')
writexl::write_xlsx(x = dt_mx_long, path = './output_student_flows_1998_2019_UN_long.xlsx')
```




