---
title: "Untitled"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---


## Set up

```{r Paths and packages}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
remove(list = ls())

### data path 
source('./Code/_path of data.R')

### packages
source('./Code/_package list.R')
library(data.table)

gc()
```

```{r Helper data}

yrs <- c(2000, 2005, 2010, 2015)

### use the Eora data as the template 
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')
load(f) # `temp_eora`

```



### Functions

```{r func_format_oda_data}
## format as the same as EORA data
func_format_oda_data <- function(data, sector_i) {
  dfs_flow_f <- data %>%
    dplyr::mutate(value = Value * 10^6) %>%   ## change the unit to USD
    rename('year' = 'Year') %>%
    # dplyr::select(year, iso3, everything()) %>%
    dplyr::filter(year %in% yrs) %>%
    dplyr::select(-c(Recipient, Donor, Value)) %>%
    ## the template can limit the years included
    merge(x = temp_eora,
          y = .,
          by.x = c('year', 'iso3', 'to'),
          by.y = c('year', 'o_iso3', 'd_iso3'), 
          all.x = T) %>%
    as.data.frame()
  
  
  ### To generate the same format as Eora in Matrix
  df_formatted <- dfs_flow_f %>%
    ungroup() %>%
    arrange(year, iso3, to) %>%
    dplyr::select(year, ctr, everything()) %>%
    spread(key = to, value = value) %>%
    arrange(year, iso3)
  
  
  ### Count NA values in the data frame ----------------------------------------------------
  cat('\n Total NA in', ind, ':\t\t', sum(is.na(df_formatted)), '\n')
  
  ### save to the `eora_cleaned` folder ----------------------------------------------------
  df_formatted %>% group_by(year) %>% tally()
  fname <- paste0(dir.eora_cleaned, 'ODA_', sector_i, '_matrix.xlsx'); print(fname)
  writexl::write_xlsx(x = df_formatted, path = fname)
}

```


## Data

### Load ODA data

  OECD offers several data sources for ODA data. 
  
  * *An overview* 
    https://www.oecd.org/dac/financing-sustainable-development/development-finance-data/
    
  * *Data* 
    - Geographical Distribution of Financial Flows to Developing Countries, a report, and also a collection of data at https://doi.org/10.1787/20743149 --> the data can be extracted and downloaded from https://stats.oecd.org/Index.aspx?datasetcode=TABLE7B# 
    
    - [Flows by Provider and Recipient](https://stats.oecd.org/Index.aspx?datasetcode=TABLE7B#) -- *BEST* data sources
      There are several options to download data. 
      To download all and do data cleaning locally, you can click Export/Related files/ and then click on the link to bulk download the tables. 
    
  * *Others*
    - Frequently asked questions on ODA https://www.oecd.org/dac/financing-sustainable-development/development-finance-data/faq.htm
    - Development finance institutions and private sector development https://www.oecd.org/development/development-finance-institutions-private-sector-development.htm 
  


  * *Unit*:	US Dollar, Millions
  
  
  `Official Flows`: the sum of Official Development Assistance (`ODA`) and Other Official Flows (`OOF`) represents the total (gross or net) disbursements by the official sector at large to the recipient country shown.

```{r - input data 1 - ODA_OOF}
###' - 1: bilateral, Total ODA, USD millions
###'      "Total Official, Gross"
###'      "Total Official,  Net"
###' TotalOFFICIAL_Data_Total official flows by country and region (ODA+OOF).zip ----------
f <- paste0(dir.raw, 'OECD_ODA/', 
            'TotalOFFICIAL_ODA_OOF/', 
            'TotalOfficial_Data_utf8.csv')
### name the indicator
ind <- 'ODA_OOF_oecd'
d <- data.table::fread(file = f, sep = ',')
unique(d$`Aid type`) ## "Total Official, Gross", "Total Official,  Net"
nrow(d) # 1,989,206

d1 <- d
```


```{r - input data 2 - ODA_disbursements}
###' - 2: bilateral, by `Aid type`, USD millions
###' Table2a_Data_Aid (ODA) disbursements to countries and regions [DAC2a].zip ------------
f <- paste0(dir.raw, 'OECD_ODA/',
            'Table2a_DAC2a/',
            'Table2a_Data_utf8.csv')
### name the indicator
ind <- 'ODA_disbursements_oecd'
d <- data.table::fread(file = f, sep = ',')
names(d)
unique(d$`Aid type`)
nrow(d) # 9,255,500

d2 <- d
```


```{r - input data 3 - ODA_DACGEO}
###' - 3: bilateral, by `Series (Aid type)`, USD millions
###' DACGEO_Data_GeoBook-Geographical flows to developing countries.zip -------------------
f <- paste0(dir.raw, 'OECD_ODA/',
            'DACGEO_Data_GeoBook-Geographical flows to developing countries/',
            'DACGEO_Data_utf8.csv')
### name the indicator
ind <- 'ODA_DACGEO_oecd'
d <- data.table::fread(file = f, sep = ',')
names(d)
unique(d$`Series`)
nrow(d) # 3,615,548

d3 <- d
```


```{r - input data 4 - ODA_DACSECTOR}
###' - 4: bilateral, by `Sector (Aid type)`, USD millions
###' DACSECTOR_Data_GeoBook ODA by sector - bilateral commitments by donor and recipient.zip
f <- paste0(dir.raw, 'OECD_ODA/',
            'DACSECTOR_Data_GeoBook ODA by sector - bilateral commitments by donor and recipient/',
            'DACSECTOR_Data_utf8.csv')
### name the indicator
ind <- 'ODA_DACSECTOR_oecd'
d <- data.table::fread(file = f, sep = ',')
names(d)
unique(d$`Sector`)
nrow(d) # 569,833

d4 <- d
```



*CRS: Creditor Reporting System (flows)*
  The objective of the CRS Aid Activity database is to provide a set of readily available granular data that enables analysis on where aid goes, what purposes it serves and what policies it aims to implement, on a comparable basis for all providers who report their activity-level statistics to the OECD. Data are collected on individual projects and programmes, with a range of attributes for each activity. Further information on the definition and possible values of each attribute can be found in the Converged Statistical Reporting Directives.

*Dimensions*: Donor, Recipient, Sector, Measure, Channel, Modality, Flow type, Price base, Drilldown dimension, Time period

*Last updated*:   November 15, 2024 at 6:29:10 AM

```{r - input data 5 - CRS }

### name the indicator
ind <- 'ODA_CRS'

fs <- list.files(path = paste0(dir.raw, 'OECD_ODA/CRS data/'), pattern = 'CRS.*txt', full.names = T)

d5 <- data.frame()

for (f in fs) {
  d <- readr::read_delim(file = f, delim ="|", show_col_types = FALSE)
  d5 <- rbind(d5, d)
}


names(d5)
# unique(d$`Sector`)
nrow(d5) # 569,833

d5 %>% distinct(FlowCode, FlowName) %>% arrange(FlowCode)
# FlowCode    FlowName
# 11	ODA Grants			
# 13	ODA Loans	
# 14	Other Official Flows (non Export Credit)
# 19	Equity Investment
# 30	Private Development Finance			
		
##' 17. Grants, loans and other flows entering the calculation of the grant equivalent ODA measure are referred to as ODA flows. 
##' 18. Other official flows (excluding officially supported export credits – see paragraph 19) are defined as transactions by the official sector which do not meet the conditions for eligibility as ODA
##' --> 11 + 13 + 14 => ODA

# Remove columns where names contain "apple"
cols_remove <- c("AgencyName", "Bi_Multi", "CrsID", "ProjectNumber", "InitialReport", "RegionName", 
                 "DonorCode", "AgencyCode", "DEIncomegroupCode",
                 "RecipientCode", 'RegionCode', 'IncomegroupCode', 'CurrencyCode',
                 "Finance_T", "Aid_T", "DERegioncode",
                 "IncomegroupName", "TypeRepayment", "NumberRepayment", "CommitmentDate" )
d5_ <- d5 %>%
  select(-matches("^PSI|^Project|^Interest|^Repay", ignore.case = F)) %>%
  select(-c(ShortDescription:SDGfocus, 
            USD_Commitment_Defl:USD_Export_Credit,
            USD_Received:Keywords,
            USD_Interest:CapitalExpend)) %>%
  select(-any_of(cols_remove)) %>%
  filter(str_detect(string = RecipientName, pattern = "unspecified", negate = T),
         str_detect(string = DERecipientcode, pattern = "_X", negate = T),
         nchar(DEDonorcode) == 3,
         nchar(DERecipientcode) == 3,
         FlowCode < 15
  ) %>%
  rename('o_iso3' = 'DEDonorcode', 
         'd_iso3' = 'DERecipientcode',
         'Donor' = 'DonorName',
         'Recipient' = 'RecipientName'
         ) %>%
  select(o_iso3, d_iso3, Recipient, Donor, Year, everything())

names(d5_)
unique(d5_$FlowName)


# 28. Biodiversity
# 29. Climate change – mitigation
# 30. Climate change – adaptation
# 31. Desertification
# # 2=principal objective
# # 1=significant objective
# # 0=not targeted
# # Blank=not screened
```


```{r - input data 5 - CRS - Biodiversity}
d5_bio <- d5_ %>%
  filter(Biodiversity > 0) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Year) %>%
  summarise_at(c("USD_Commitment", "USD_Disbursement"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()


d5_bio %>%
  ggplot() +
  geom_point(aes(x=USD_Commitment, y=USD_Disbursement)) +
  geom_abline(slope = 1, color = 'red') +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))


d5_Biodiversity <- d5_bio %>%
  select(-USD_Commitment) %>%
  rename('Value' = 'USD_Disbursement')

## format data and save as xlsx
func_format_oda_data(data = d5_Biodiversity, sector_i = 'Biodiversity')
```


### Country names and code

```{r - tidy names}
### clean the country names and add iso3 code --------------------------------------------
f <- paste0(dir.raw, 'OECD_ODA/', "Recipient_13_266.xlsx")

# List all the sheets in the Excel file
sheet_names <- readxl::excel_sheets(f)
sheet_names


### test code using one sheet ------------------------------------
# c.d <- readxl::read_excel(f, sheet = 'ODA_OOF_Donor') %>% as.data.frame()
#   
# c.d_cleaned <- as.data.frame(lapply(c.d, str_squish)) %>%
#   # dplyr::filter(str_detect(c3,'Information')) %>%
#   dplyr::mutate(across(where(is.character), ~ str_replace_all(., "Information", ""))) %>%
#   dplyr::mutate(across(where(is.character), str_squish))


### loop all sheets and clean the country names -------------------
c_all <- data.frame()
for (sheet in sheet_names) {
  c <- readxl::read_excel(f, sheet = sheet) %>% as.data.frame()

  # Remove extra spaces from all character columns using str_squish()
  c_cleaned <- as.data.frame(lapply(c, str_squish)) %>%
    # dplyr::filter(str_detect(c3,'Information')) %>%
    dplyr::mutate(across(where(is.character), ~ str_replace_all(., "Information", ""))) %>%
    dplyr::mutate(across(where(is.character), str_squish))
  
  ## 
  c_all <- rbind(c_all, c_cleaned) %>%
    dplyr::distinct_all(.keep_all = T) %>%
    arrange(c3, c2, c1)
}


### a list of institutes that are not countries
c.exlcude <- c(c_all$c1, c_all$c2) %>% 
  unique() %>% 
  sort()



### extended country-iso3 pair list for matching
load('./Data/Ancillary_Data_iso_eora_ex.RData')  ## ## iso_eora_ex
iso3_code <- iso_eora_ex %>% 
  dplyr::select(-IMF) %>%
  dplyr::distinct_all() %>%
  as.data.frame()

c.ctr <- c_all %>% 
  dplyr::distinct(c3) %>%
  dplyr::filter(str_detect(
    c3, 
    paste('region', 'Unspecified', 'unspecified', 'Community', 'Multilateral', 'Project', 
          'Other donor', 'Indus Basin', 'EU Institutions', sep = '|'), 
    negate = T)) %>%
  # dplyr::rename('name' = 'c3') %>%
  dplyr::mutate(
    name_clean = str_replace(c3, " \\s*\\([^\\)]+\\)", "")
  ) %>%
  left_join(
    x = .,
    y = iso3_code,
    by = c('name_clean' = 'Row')
  ) %>%
  arrange(!is.na(iso3_eora)) %>%
  dplyr::select(-name_clean) %>%
  as.data.frame()

rm(c, c_cleaned, c_all)
```



### Tidy data

```{r function}

func_join_country_code <- function(data) {
  d <- data %>%
    ## remove non-country names
    dplyr::filter(!Recipient %in% c(
      c.exlcude, 'CABEI', 'EC', 'EDF', 'EBRD', 'EIB', 'CEC FOR CEEC/NIS',
      'IBRD', 'IDB', 'MIGA', 'IFC', 'IDA-MDRI', 'IDA', 'IMF',
      'IDB Special Oper. Fund', 'Caribbean Dev. Bank', 'Af. D B', 'African Dev. Fund',
      'Other EU institutions', 'Other UN',
      'Asian Dev. Bank'),
                  !Donor %in% c.exlcude) %>%
    dplyr::filter(str_detect(Donor, 'Agricole Foundation', negate = T)) %>%
    dplyr::filter(str_detect(Recipient, 'Unspecified|Multilateral|Other Regional Banks|Part II|Other World Bank|Special Funds', negate = T)) %>%
    dplyr::mutate(
        Donor = ifelse(str_detect(Donor, "Ivoire"), "Côte d'Ivoire", Donor),
        Donor = ifelse(str_detect(Donor, "kiye"), "Türkiye", Donor),
        Recipient = ifelse(str_detect(Recipient, "Ivoire"), "Côte d'Ivoire", Recipient),
        Recipient = ifelse(str_detect(Recipient, "kiye"), "Türkiye", Recipient),
        ) %>%
    
    ## add country code
    left_join(x = .,
              y = c.ctr, 
              by = c('Recipient' = 'c3')) %>%
    dplyr::rename('d_iso3' = 'iso3_eora') %>%
    left_join(x = .,
              y = c.ctr, 
              by = c('Donor' = 'c3')) %>%
    dplyr::rename('o_iso3' = 'iso3_eora') %>%
    arrange(!is.na(o_iso3), !is.na(d_iso3)) %>%
    as.data.frame()
  
  return(d)
}


```


```{r}
# The 'fread' function automatically detects the delimiter (comma in this case)

cols_to_remove <- c("RECIPIENT", "DONOR", "DATATYPE",
                    "AIDTYPE", "SERIES",
                    "SECTOR", 
                    "PART", 'AMOUNTTYPE', 'TIME', 'Flags')


data <- d3

names(data)
# unique(data$Recipient)
# unique(data$Donor)
unique(data$`Aid type`)
unique(data$Series)
unique(data$Sector)

unique(data$Part)          ## "1 : Part I - Developing Countries"     "2 : Part II - Countries in Transition"
unique(data$`Amount type`) ## "Current Prices (USD millions)"       "Constant Prices (2021 USD millions)"
unique(data$Sector) %>% str_to_title()
```

#### data 1 - Total Official, Gross
```{r}
d1_total_o_gross <- d1 %>%
  filter(`Aid type` == "Total Official, Gross",
         `Amount type` == "Constant Prices (2021 USD millions)") %>%
  dplyr::filter(Year >= 2000) %>%
  ## these columns represent but repeat the text columns 
  dplyr::select(-any_of(cols_to_remove)) %>%
  func_join_country_code(data = .) %>%
  as.data.frame() 


unique(d1_total_o_gross$Part)
names(d1_total_o_gross)

d1_total_o_gross %>%
  distinct(o_iso3, d_iso3, Year) %>%
  nrow()


d1_total_o_gross <- d1_total_o_gross %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()


## format data and save as xlsx
func_format_oda_data(data = d1_total_o_gross, sector_i = 'OOF_total') ## Total Official = ODA + OOF
```




#### data 2 - `ODA Total, Gross disbursements`; `ODA: Total Net`

```{r}

d2_ <- d2 %>%
  dplyr::filter(Year >= 2000) %>%
  ## these columns represent but repeat the text columns 
  dplyr::select(-any_of(cols_to_remove)) %>%
  func_join_country_code(data = .) %>%
  filter(`Amount type` == "Constant Prices (2021 USD millions)") %>%
  as.data.frame() 

names(d2_)
unique(d2_$`Aid type`) %>% sort()


d2_total_oda_gross <- d2_ %>%
  filter(`Aid type` == "Memo: ODA Total, Gross disbursements") %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()
## format data and save as xlsx
func_format_oda_data(data = d2_total_oda_gross, sector_i = 'total_disbursements')
```





#### data 3 - Total Official Gross, Total ODA Gross, Food Aid

```{r}
d3_ <- d3 %>%
  ### rename a column to keep consistent 
  setNames(gsub("Series", "Aid type", names(.))) %>%
  ### Adding column if it does not exist (because `DACGEO_oecd`does not have this column)
  rowwise() %>%
  dplyr::mutate(
    `Amount type` = ifelse("Amount type" %in% names(.), 
                           `Amount type`, 
                           "Constant Prices (2021 USD millions)")) %>%
  dplyr::filter(Year >= 2000) %>%
  ## these columns represent but repeat the text columns 
  dplyr::select(-any_of(cols_to_remove)) %>%
  as.data.frame() 


d3_sec <- d3_ %>%
  dplyr::filter(
    `Aid type` %in% c("Food Aid",             # d3
                      "Total OOF Gross",      # d3
                      "Total Official Gross", # d3
                      "Total ODA Gross",      # d3
                      "Total ODA Net"),       # d3
    `Amount type` == "Constant Prices (2021 USD millions)",
    Year >= 2000
  ) %>%
  func_join_country_code(data = .) %>%
  as.data.frame()

unique(d3_sec$`Aid type`)



### if there are > 1 `Aid type` in the data, we need to further decide which one to keep
d3_sec_check <- d3_sec %>%
  spread(key = 'Aid type', value = 'Value')

d3_sec_check %>%
  # ggplot(aes(x = `ODA Gross Loans`, y = `ODA: Total Net`)) + ## `d2`
  ggplot(aes(x = `Total ODA Gross`, y = `Total Official Gross`)) +  ## `d3`
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, color = 'red') +
  labs(x = 'Total ODA Gross', y = 'Total Official Gross') +
  theme_bw()

# ## check number of NA
# ##' for `d3`
# sum(is.na(d3_sec_check$`Total ODA Gross`)) ## larger values
# sum(is.na(d3_sec_check$`Total Official Gross`))   ## less NA
```


```{r - Total Official Gross}

### aggregate data after filtering -------------------------------------------------------
d3_total_o_gross <- d3_sec %>%
  dplyr::filter(`Aid type` == "Total Official Gross") %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, `Aid type`, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()


d3_total_oda_gross <- d3_sec %>%
  dplyr::filter(`Aid type` == "Total ODA Gross") %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()
# ## format data and save as xlsx
# func_format_oda_data(data = d3_total_oda_gross, sector_i = 'total')
# ##' ! `d2_total_oda_gross` is better with fewer NA


## ! no country specific data for this indicator
d3_total_oof_gross <- d3_sec %>%
  dplyr::filter(`Aid type` == "Total OOF Gross") %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, `Aid type`, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()

```



```{r - Food aid}
sector_i <- 'Food' 
d3_sec_food <- d3_sec %>%
  dplyr::filter(stringr::str_detect(`Aid type`, regex(sector_i, ignore_case=TRUE))) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, `Aid type`, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()
```



####  data 4 - by sector

  Refer to `./Data/data_01_raw/OECD_ODA/ODA_data_compare.xlsx`, which is derived from [OECD](https://data-explorer.oecd.org/). 

```{r}
d4_sec <- d4 %>%
    ### rename a column to keep consistent 
    # setNames(gsub("Series", "Aid type", names(.))) %>%
  dplyr::select(-any_of(cols_to_remove)) %>%
  dplyr::filter(
    Year >= 2000
  ) %>%
  as.data.frame() 
names(d4_sec)

d4_sec_agg <- d4_sec %>%
  group_by(Recipient, Donor, Sector, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  func_join_country_code(data = .) %>%
  as.data.frame()
```


```{r - data by sector}
sector_i <- 'Agriculture'
d4_sec_agriculture <- d4_sec_agg %>%
  dplyr::filter(stringr::str_detect(Sector, regex(sector_i, ignore_case=TRUE))) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Sector, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  # dplyr::filter(Year == 2015) %>%
  select(-Sector) %>%
  as.data.frame()

## format data and save as xlsx
func_format_oda_data(data = d4_sec_agriculture, sector_i = sector_i)



sector_i <- 'Energy'
d4_sec_energy <- d4_sec_agg %>%
  dplyr::filter(stringr::str_detect(Sector, regex(sector_i, ignore_case=TRUE))) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Sector, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  # dplyr::filter(Year == 2015) %>%
  select(-Sector) %>%
  as.data.frame()
## format data and save as xlsx
func_format_oda_data(data = d4_sec_agriculture, sector_i = sector_i)




sector_i <- 'Education'
d4_sec_education <- d4_sec_agg %>%
  dplyr::filter(stringr::str_detect(Sector, regex(sector_i, ignore_case=TRUE))) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Sector, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  # dplyr::filter(Year == 2015) %>%
  select(-Sector) %>%
  as.data.frame()
## format data and save as xlsx
func_format_oda_data(data = d4_sec_agriculture, sector_i = sector_i)





sector_i <- 'Food'
d4_sec_food <- d4_sec_agg %>%
  dplyr::filter(stringr::str_detect(Sector, regex(sector_i, ignore_case=TRUE))) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Sector, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  # dplyr::filter(Year == 2015) %>%
  select(-Sector) %>%
  as.data.frame()
## format data and save as xlsx
func_format_oda_data(data = d4_sec_agriculture, sector_i = sector_i)



sector_i <- 'Infrastructure'
d4_sec_infr <- d4_sec_agg %>%
  dplyr::filter(stringr::str_detect(Sector, regex('Infrastructure|Transport', ignore_case=TRUE))) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  # dplyr::filter(Year == 2015) %>%
  as.data.frame()
# unique(d4_sec_agg_infr$Sector)
## format data and save as xlsx
func_format_oda_data(data = d4_sec_agriculture, sector_i = sector_i)



sector_i <- 'Water'
d4_sec_water <- d4_sec_agg %>%
  dplyr::filter(stringr::str_detect(Sector, regex(sector_i, ignore_case=TRUE))) %>%
  group_by(o_iso3, d_iso3, Recipient, Donor, Sector, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  # dplyr::filter(Year == 2015) %>%
  select(-Sector) %>%
  as.data.frame()
## format data and save as xlsx
func_format_oda_data(data = d4_sec_agriculture, sector_i = sector_i)
```



#### compare data

```{r - o_gross}

##' `d1_total_o_gross` VS `d3_total_o_gross` 
names(d1_total_o_gross)
names(d3_total_o_gross)

dd <- d1_total_o_gross %>%
  merge(x = ., 
        y = d3_total_o_gross, 
        all = T, 
        by = c("o_iso3", "d_iso3", "Donor", "Recipient", "Year" ))
  
dd %>%
  ggplot() +
  geom_point(aes(x = Value.x, y = Value.y)) +
  geom_abline(slope = 1, color = 'red')
```



```{r - oda_gross}

## `d2_total_oda_gross` VS `d3_total_oda_gross`
names(d2_total_oda_gross)
names(d3_total_oda_gross)

dd <- d2_total_oda_gross %>%
  merge(x = ., 
        y = d3_total_oda_gross, 
        all = T, 
        by = c("o_iso3", "d_iso3", "Donor", "Recipient", "Year" ))
  
dd %>%
  ggplot() +
  geom_point(aes(x = Value.x, y = Value.y)) +
  geom_abline(slope = 1, color = 'red')
```


```{r - food}
## compare sector data -- food -----------------------------------------------------------
names(d3_sec_food)
names(d4_sec_food)

dd <- d3_sec_food %>%
  merge(x = ., 
        y = d4_sec_food, 
        all = T, 
        by = c("o_iso3", "d_iso3", "Donor", "Recipient" , "Year" ))
  
dd %>%
  ggplot() +
  geom_point(aes(x = Value.x, y = Value.y)) +
  geom_abline(slope = 1, color = 'red')

```






### Final data

```{r - format the same as EORA}

##' Load function `func_format_oda_data`

```


```{r - compare}
f <- paste0(dir.eora_cleaned, 'ODA_OOF_total_matrix', '.xlsx')
df1 <- readxl::read_excel(f) %>%
  dplyr::filter(year == 2015) %>%
  gather(key = 'to', value = 'value', 4:ncol(.))

f <- paste0(dir.eora_cleaned, 'ODA_Infrastructure_matrix', '.xlsx')
df4 <- readxl::read_excel(f)%>%
  dplyr::filter(year == 2015) %>%
  gather(key = 'to', value = 'value', 4:ncol(.))

df14 <- df1 %>%
  left_join(., df4, by=c("year","ctr","iso3", "to"))

df14 %>%
  ggplot(aes(x = value.x, y = value.y)) +
  geom_point() +
  labs(x = 'df1', y = 'df4') +
  geom_abline(slope = 1, color = 'red') +
  # scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #               labels = trans_format("log10", math_format(10^.x))) +
  # scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #               labels = trans_format("log10", math_format(10^.x))) +
  theme_bw()
```

