---
title: "Untitled"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---


# Set up
```{r Paths and packages}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
remove(list = ls())

### data path 
source('./Code/_path of data.R')

### packages
source('./Code/_package list.R')

gc()
```



## Data

### - ODA OECD

  OECD offers several data sources for ODA data. 
  
  * *An overview* 
    https://www.oecd.org/dac/financing-sustainable-development/development-finance-data/
    
  * *Data* 
    - Geographical Distribution of Financial Flows to Developing Countries, a report, and also a collection of data at https://doi.org/10.1787/20743149 --> the data can be extracted and downloaded from https://stats.oecd.org/Index.aspx?datasetcode=TABLE7B# 
    
    - [Flows by Provider and Recipient](https://stats.oecd.org/Index.aspx?datasetcode=TABLE7B#) -- *BEST* data sources
      There are several options to download data. 
      To download all and do data cleaning locally, you can click Export/Related files/ and then click on the link to bulk download the tables. 
    
  * *Others*
    - Frequently asked questions on ODA https://www.oecd.org/dac/financing-sustainable-development/development-finance-data/faq.htm
    - Development finance institutions and private sector development https://www.oecd.org/development/development-finance-institutions-private-sector-development.htm 
  


  Unit	US Dollar, Millions

```{r - input data- which ? }

###' - 1
###' TotalOFFICIAL_Data_Total official flows by country and region (ODA+OOF).zip ----------
f <- paste0(dir.raw, 'OECD_ODA/', 
            'TotalOFFICIAL_ODA_OOF/', 
            'TotalOfficial_Data_utf8.csv')
### name the indicator
ind <- 'ODA_OOF_oecd'



###' - 2
###' Table2a_Data_Aid (ODA) disbursements to countries and regions [DAC2a].zip ------------
# f <- paste0(dir.raw, 'OECD_ODA/',
#             'Table2a_DAC2a/',
#             'Table2a_Data_utf8.csv')
# ### name the indicator
# ind <- 'ODA_disbursements_oecd'



###' - 3
###' DACGEO_Data_GeoBook-Geographical flows to developing countries.zip -------------------
# f <- paste0(dir.raw, 'OECD_ODA/', 
#             'DACGEO_Data_GeoBook-Geographical flows to developing countries/', 
#             'DACGEO_Data_utf8.csv')
# ### name the indicator
# ind <- 'ODA_DACGEO_oecd'



###' - 4
###' DACSECTOR_Data_GeoBook ODA by sector - bilateral commitments by donor and recipient.zip
f <- paste0(dir.raw, 'OECD_ODA/',
            'DACSECTOR_Data_GeoBook ODA by sector - bilateral commitments by donor and recipient/',
            'DACSECTOR_Data_utf8.csv')
### name the indicator
ind <- 'ODA_DACSECTOR_oecd'
```



#### * Tidy data

```{r}
# The 'fread' function automatically detects the delimiter (comma in this case)

cols_to_remove <- c("RECIPIENT", "DONOR", 
                    "AIDTYPE", "SERIES",
                    "SECTOR", 
                    "PART", 'AMOUNTTYPE', 'TIME', 'Flags')

library(data.table)
data <- data.table::fread(file = f, sep = ',')

names(data)
# unique(data$Recipient)
# unique(data$Donor)
unique(data$`Aid type`)
unique(data$Series)


unique(data$Part)          ## "1 : Part I - Developing Countries"     "2 : Part II - Countries in Transition"
unique(data$`Amount type`) ## "Current Prices (USD millions)"       "Constant Prices (2021 USD millions)"
unique(data$Sector) %>% str_to_title()
```



```{r - data 1, eval=FALSE, include=FALSE}
if (ind == 'ODA_DACGEO_oecd'){
  dat_ <- data %>%
    ### rename a column to keep consistent 
    setNames(gsub("Series", "Aid type", names(.))) %>%
    ### Adding column if it does not exist (because `DACGEO_oecd`does not have this column)
    rowwise() %>%
    dplyr::mutate(
      `Amount type` = ifelse("Amount type" %in% names(.), 
                             `Amount type`, 
                             "Constant Prices (2021 USD millions)")) %>%
    as.data.frame() 
} else if (ind != 'ODA_DACSECTOR_oecd') {
  dat_ <- data
} else {
  cat('\n', 'please load the `by sector` data', '\n\n')
}


data_ <- dat_ %>%
  dplyr::select(-any_of(cols_to_remove)) %>%
  dplyr::filter(
    `Aid type` %in% c("Total Official, Gross", 
                      "ODA Gross Loans", 
                      "ODA: Total Net",
                      "Total ODA Net"),
    `Amount type` == "Constant Prices (2021 USD millions)",
    Year >= 2010
  )

unique(data_$`Aid type`)

### if there are > 1 `Aid type` in the data, we need to further decide which one to keep
if( length(unique(data_$`Aid type`)) > 1) {
  data_check <- data_ %>%
    spread(key = 'Aid type', value = 'Value')
  
  data_check %>%
    ggplot(aes(x = `ODA Gross Loans`, y = `ODA: Total Net`)) +
    geom_point() +
    geom_abline(slope = 1) +
    theme_bw()
  
  ## check number of NA
  sum(is.na(data_check$`ODA Gross Loans`))
  sum(is.na(data_check$`ODA: Total Net`))  ## less NA, large values
  
  ## keep one of the 'Aid type' for further analysis
  data_aidtype <- data_ %>%
    dplyr::filter(`Aid type` == "ODA: Total Net")
  unique(data_aidtype$`Aid type`)  
} else {
  data_aidtype <- data_
}


### aggregate data after filtering -------------------------------------------------------
cols_to_remove2 <- c("Aid type", "Amount type", "Part")

data_agg <- data_aidtype %>%
  dplyr::select(-any_of(cols_to_remove2)) %>%
  group_by(Recipient, Donor, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()

data_wide <- data_agg %>%
  spread(key = 'Recipient', value = 'Value')


names(data_agg)
rm(data)
gc()
```




```{r - data 2 - by sector, eval=FALSE, include=FALSE}

data_sec <- data %>%
    ### rename a column to keep consistent 
    # setNames(gsub("Series", "Aid type", names(.))) %>%
  dplyr::select(-any_of(cols_to_remove)) %>%
  dplyr::filter(
    Year >= 2010
  ) %>%
  as.data.frame() 


data_sec_agg <- data_sec %>%
  group_by(Recipient, Donor, Sector, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  as.data.frame()


data_sec_agg_infr <- data_sec_agg %>%
  dplyr::filter(stringr::str_detect(Sector, regex('Infrastructure|Multisector|Transport', ignore_case=TRUE))) %>%
  group_by(Recipient, Donor, Year) %>%
  dplyr::summarise_at(c("Value"), sum, na.rm = TRUE) %>%
  ungroup() %>%
  # dplyr::filter(Year == 2015) %>%
  as.data.frame()

# unique(data_sec_agg_infr$Sector)
```



```{r - compare data}
if (ind == 'ODA_DACSECTOR_oecd'){
  data_agg <- data_sec_agg_infr %>%
    dplyr::mutate(
      Donor = ifelse(str_detect(Donor, "Ivoire"), "Côte d'Ivoire", Donor),
      Donor = ifelse(str_detect(Donor, "rkiye"), "Türkiye", Donor),
      Recipient = ifelse(str_detect(Recipient, "Ivoire"), "Côte d'Ivoire", Recipient),
      Recipient = ifelse(str_detect(Recipient, "rkiye"), "Türkiye", Recipient),
      ) %>%
    as.data.frame()
} else{
  data_agg = data_agg
}


### 
cat('\n Total rows of data in', ind, ': \t\t', nrow(data_agg), '\n')

##' Total rows of data in ODA_OOF_oecd : 		         151609
##' Total rows of data in ODA_disbursements_oecd : 	 164060 -- best?
##' Total rows of data in DACGEO_oecd : 		         126893 
##' Total rows of data in ODA_DACSECTOR_oecd : 		    46758
```


#### * Country names
```{r - tidy names}
### clean the country names and add iso3 code --------------------------------------------
f <- paste0(dir.raw, 'OECD_ODA/', "Recipient_13_266.xlsx")

# List all the sheets in the Excel file
sheet_names <- readxl::excel_sheets(f)
sheet_names


### test code using one sheet ------------------------------------
# c.d <- readxl::read_excel(f, sheet = 'ODA_OOF_Donor') %>% as.data.frame()
#   
# c.d_cleaned <- as.data.frame(lapply(c.d, str_squish)) %>%
#   # dplyr::filter(str_detect(c3,'Information')) %>%
#   dplyr::mutate(across(where(is.character), ~ str_replace_all(., "Information", ""))) %>%
#   dplyr::mutate(across(where(is.character), str_squish))


### loop all sheets and clean the country names -------------------
c_all <- data.frame()
for (sheet in sheet_names) {
  c <- readxl::read_excel(f, sheet = sheet) %>% as.data.frame()

  # Remove extra spaces from all character columns using str_squish()
  c_cleaned <- as.data.frame(lapply(c, str_squish)) %>%
    # dplyr::filter(str_detect(c3,'Information')) %>%
    dplyr::mutate(across(where(is.character), ~ str_replace_all(., "Information", ""))) %>%
    dplyr::mutate(across(where(is.character), str_squish))
  
  ## 
  c_all <- rbind(c_all, c_cleaned) %>%
    dplyr::distinct_all(.keep_all = T) %>%
    arrange(c3, c2, c1)
}


### a list of institutes that are not countries
c.exlcude <- c(c_all$c1, c_all$c2) %>% 
  unique() %>% 
  sort()



### extended country-iso3 pair list for matching
load('./Data/Ancillary_Data_iso_eora_ex.RData')  ## ## iso_eora_ex
iso3_code <- iso_eora_ex %>% 
  dplyr::select(-IMF) %>%
  dplyr::distinct_all() %>%
  as.data.frame()

c.ctr <- c_all %>% 
  dplyr::distinct(c3) %>%
  dplyr::filter(str_detect(
    c3, 
    paste('region', 'Unspecified', 'unspecified', 'Community', 'Multilateral', 'Project', 
          'Other donor', 'Indus Basin', 'EU Institutions', sep = '|'), 
    negate = T)) %>%
  # dplyr::rename('name' = 'c3') %>%
  dplyr::mutate(
    name_clean = str_replace(c3, " \\s*\\([^\\)]+\\)", "")
  ) %>%
  left_join(
    x = .,
    y = iso3_code,
    by = c('name_clean' = 'Row')
  ) %>%
  arrange(!is.na(iso3_eora)) %>%
  dplyr::select(-name_clean) %>%
  as.data.frame()

rm(c, c_cleaned, c_all)
```



#### * Final data

```{r - filter and tidy data}
data_ctr <- data_agg %>%
  dplyr::filter(!Recipient %in% c.exlcude,
                !Donor %in% c.exlcude) %>%
  left_join(x = .,
            y = c.ctr, 
            by = c('Recipient' = 'c3')) %>%
  dplyr::rename('iso3_reci' = 'iso3_eora') %>%
  left_join(x = .,
            y = c.ctr, 
            by = c('Donor' = 'c3')) %>%
  dplyr::rename('iso3_send' = 'iso3_eora') %>%
  arrange(!is.na(iso3_send), !is.na(iso3_reci)) %>%
  as.data.frame()



# c.check <- data_ctr %>%
#   dplyr::distinct(Recipient, Donor) %>%
#   # dplyr::distinct_all() %>%
#   # dplyr::distinct_at(vars(Recipient, Donor)) %>%
#   as.data.frame()
# 
# c.check.r <- c.check %>%
#   dplyr::distinct(Recipient)
# 
# c.check.d <- c.check %>%
#   dplyr::distinct(Donor)
```




```{r - format the same as EORA}

yrs <- c(2000, 2005, 2010, 2015)

### use the Eora data as the template 
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')
load(f) # `temp_eora`


## format as the same as EORA data
dfs_flow_f <- data_ctr %>%
  dplyr::mutate(value = Value * 10^6) %>%   ## change the unit to USD
  rename('year' = 'Year') %>%
  # dplyr::select(year, iso3, everything()) %>%
  dplyr::filter(year %in% yrs) %>%
  dplyr::select(-c(Recipient, Donor, Value)) %>%
  ## the template can limit the years included
  merge(x = temp_eora,
        y = .,
        by.x = c('year', 'iso3', 'to'),
        by.y = c('year', 'iso3_send', 'iso3_reci'), all.x = T) %>%
  as.data.frame()



### To generate the same format as Eora in Matrix
df_formatted <- dfs_flow_f %>%
  ungroup() %>%
  arrange(year, iso3, to) %>%
  dplyr::select(year, ctr, everything()) %>%
  spread(key = to, value = value) %>%
  arrange(year, iso3)


### Count NA values in the data frame ----------------------------------------------------
cat('\n Total NA in', ind, ':\t\t', sum(is.na(df_formatted)), '\n')

##' Total NA in ODA_OOF_oecd :		           136306 -- best
##' Total NA in ODA_disbursements_oecd :		 136332
##' Total NA in DACGEO_oecd :		             136594
##' 
df_formatted.check <- df_formatted %>%
  dplyr::filter(year == 2015) %>%
  gather(key = 'Recipient', value = 'value', 4:ncol(.)) 

df_formatted.check2 <- df_formatted.check %>%
  dplyr::filter(year == 2015, 
                ctr == 'Australia') %>%
  dplyr::mutate(value_million = value/10^6)
cat('\n Total NA in', ind, 'in 2015:\t\t', sum(is.na(df_formatted.check$value)), '\n')

##' Total NA in ODA_OOF_oecd in 2015:		           32351 -- best
##' Total NA in ODA_disbursements_oecd in 2015:		 32368
##' Total NA in DACGEO_oecd in 2015:		           32469 
##' 
##' It seems that *ODA_OOF_oecd* is a better data source!
##' 


### save to the `eora_cleaned` folder ----------------------------------------------------
df_formatted %>%
  group_by(year) %>% tally()
fname <- paste0(dir.eora_cleaned, ind, '_matrix.xlsx'); fname
writexl::write_xlsx(x = df_formatted, path = fname)
gc()
```

