---
title: "FDI"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---


# Set up
```{r Paths and packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
remove(list = ls())

### set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
getwd()

### data path 
source('./Code/_path of data.R')


### packages
source('./Code/_package list.R')
```


# Data

## Ancillary data
```{r}

### extended country-iso3 pair list for matching
load('./Data/Ancillary_Data_iso_eora_ex.RData')  ## ## iso_eora_ex


### use the Eora data as the template 
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')
load(f) # `temp_eora`
```



## FDI
  Based on the list by Harvard - *Foreign Direct Investment and Aid Data Resources* at https://guides.library.harvard.edu/fdi, this data -- *Coordinated Direct Investment Survey - IMF* is what we need. 
  
  Link to the data: https://data.imf.org/?sk=40313609-f037-48c1-84b1-e1f1ce54d6d5&sid=1410469360660 --> *CDIS Table 6: Direct lnvestment Positions by All Reporting Economies Cross-classified by Counterpart Economies*

  *Meta-data*  
  * Blank cells reflect data not available or not applicable and cells with "c" reflect data that were suppressed by the reporting economy to preserve confidentiality. “0” reflects amounts that are less than +/-$500,000, or amounts reported as “0”.
  Totals may not be equal to the sum of their components due to rounding. 
  
  Based on `Metadata availability` https://data.imf.org/regular.aspx?key=60559496, *Inward* data are better. 
  
  *Other related info*
  - https://data.oecd.org/fdi/fdi-flows.htm#indicator-chart 
  - [cepii - 2012 old data](http://www.cepii.fr/CEPII/en/bdd_modele/bdd_modele_item.asp?id=4)

```{r - IMF}
### get the indicator name
ind <- 'FDI'

yrs <- c(2000, 2005, 2010, 2015)
yrs <- c(2000, 2005, 2010, 2015, 2021)

dir.i <- './Data/data_01_raw/FDI/IMF/'

fdi.direction <- "inward";  column1.name  <- "Investment from:"
# fdi.direction <- "outward"; column1.name  <- "Investment in:"


file.pat <- paste0('CDIS_Table_6_Direct_Investment_Posi-', fdi.direction)
xls_ls <- list.files(path = dir.i, pattern = file.pat, full.names = T); xls_ls


dfs <- data.frame()
dfs_flow <- data.frame()

## test
# i = 1
# xls <- xls_ls[i]

for (xls in xls_ls) {
  
  yr <- readxl::read_excel(path = xls, range = "A2") 
  yr <- stringr::str_sub(string = names(yr), start = -4, end = -1)
  cat(xls, '\n\t year:', yr, '\n')
  
  df <- readxl::read_excel(path = xls, skip = 3, n_max = 248) %>%
    dplyr::mutate_all(., as.character) %>%
    dplyr::rename("column1" = column1.name) %>%
    as.data.frame()
  
  ## to check if the last row of data is correct, otherwise, we need to change `n_max` above
  value_check <- df$column1[nrow(df)]
  
  if(value_check != "World") {
    stop("Invalid value detected")
  }
   
  
  ## clean the data 
  ## 
  iso3_code <- iso_eora_ex %>% dplyr::select(-IMF)
  df_ <- df %>%
    dplyr::filter(str_detect(pattern = 'Not Specified', negate = T, string = column1)) %>%
    pivot_longer(cols = 2:ncol(.), names_to = 'cols', values_to = 'USD_million') %>%
    dplyr::mutate(USD_million = gsub('C', NA, USD_million),
                  USD_million = as.numeric(USD_million)) %>%
    
    ### add country code ISO3
    ### 1. for `from` ---------------------
    # dplyr::distinct(column1, .keep_all = T) %>% ## for easy-checking
    dplyr::mutate(column1 = gsub('\\.|\\:|\\,', '', column1),
                  column1 = trimws(column1),
                  column1 = str_squish(column1)) %>%
    left_join(x = .,
              y = iso3_code,
              by=c('column1' = 'Row')) %>%
    dplyr::rename("iso3" = "iso3_eora") %>%
    arrange(!is.na(iso3)) %>%
    
    ### 2. for `to` ---------------------
    # dplyr::distinct(cols, .keep_all = T) %>%  ## for easy-checking
    dplyr::mutate(cols = gsub('\\.|\\:|\\,', '', cols),
                  cols = trimws(cols),
                  cols = str_squish(cols)) %>%
    left_join(x = .,
              y = iso3_code,
              by=c('cols' = 'Row')) %>%
    dplyr::rename("iso3_cols" = "iso3_eora") %>%
    arrange(!is.na(iso3_cols)) %>%
    
    dplyr::mutate(year = as.numeric(yr)) %>%
    as.data.frame()
  
  
  df_flow <- df_ %>%
    dplyr::filter(column1 != 'World') %>%
    dplyr::filter(cols    != 'Total Investment') %>%
    as.data.frame()
  
  
  cat('\t from is NA:', sum(is.na(df_flow$iso3)), '\n',
      '\t   to is NA:', sum(is.na(df_flow$iso3_cols)),   '\n')
  cat('\t--- done! ---\n\n')
  
  ## rbind data -------------------------
  dfs <- rbind(dfs, df_)
  dfs_flow <- rbind(dfs_flow, df_flow)

}
  

## label flow direction 
if (fdi.direction == 'inward') {
  dfs_flow_ <- dfs_flow %>%
    dplyr::select(-column1, -cols) %>% ## remove the columns with country's full name (column1 = from, cols = to)
    dplyr::rename(to    = iso3_cols,   ## 'iso3_cols' is the receiving countries 
                  value = USD_million)
} else {
  dfs_flow_ <- dfs_flow %>%
    dplyr::select(-column1, -cols) %>% ## remove the columns with country's full name (column1 = to, cols = from)
    dplyr::rename(to    = iso3,        ## 'iso3' is the receiving countries 
                  iso3  = iso3_cols,   ## 'iso3_cols' is the sending countries 
                  value = USD_million)
}


### save IMF data to the raw data folder
fname <- paste0(dir.i, ind, '_rawCleaned_', fdi.direction,'.xlsx'); fname
writexl::write_xlsx(x = dfs_flow_, path = fname)
```


  By looking at Table 3 - https://data.imf.org/regular.aspx?key=60564262, we can also spot the different values reported by either country side. 
  
```{r - IMF compare inward vs outward}
yr_comp <- c(2010, 2015, 2021)
yr_comp <- c(2021)

f <- paste0(dir.i, 'FDI_rawCleaned_inward.xlsx');
d1 <- readxl::read_excel(path = f) %>%
  # pivot_longer(cols = 4:ncol(.), names_to = 'to', values_to = 'value_in') %>%
  # dplyr::select(-ctr) %>%
  # dplyr::filter(year %in% yr_comp) %>%
  dplyr::rename('value_in' = 'value') %>%
  as.data.frame()
  
  

f <- paste0(dir.i, 'FDI_rawCleaned_outward.xlsx');
d2 <- readxl::read_excel(path = f) %>%
  # pivot_longer(cols = 4:ncol(.), names_to = 'to', values_to = 'value_out')%>%
  # dplyr::select(-ctr) %>%
  # dplyr::filter(year %in% yr_comp) %>%
  dplyr::rename('value_out' = 'value') %>%
  as.data.frame()

dcomp <- d1 %>%
  left_join(., d2, 
            by = c("year","iso3","to")) %>%
  dplyr::filter(year %in% yr_comp) %>%
  dplyr::select(year, iso3, to, everything())

dcomp %>%
  ggplot() +
  geom_point(aes(x= value_in, y= value_out)) +
  geom_abline(slope=1)

dcomp %>%
  dplyr::filter(iso3 == 'USA', to == 'AUS')

dcomp %>%
  dplyr::filter(iso3 == 'AUS', to == 'USA')
```



  Using both outward data to fill NA in `inward`, and use the average value if both are available. 
  
```{r - IMF - final 📁}
## use both data to further clean the dataset 
dcomp_ <- d1 %>%
  left_join(., d2, 
            by = c("year","iso3","to")) %>%
  dplyr::select(year, iso3, to, everything()) %>%
  dplyr::mutate(value = case_when(
    !is.na(value_in) & is.na(value_out)  ~ value_in,
    is.na(value_in) & !is.na(value_out)  ~ value_out,
    !is.na(value_in) & !is.na(value_out) ~ (value_in + value_out)/2,
    T ~ NA
  )) %>%
  dplyr::select(-value_in,-value_out)


## format as the same as EORA data
dfs_flow_f <- dcomp_ %>%
  dplyr::mutate(value = value * 10^6) %>%   ## change the unit to USD
  dplyr::select(year, iso3, everything()) %>%
  
  ## use data in 2009 to represent data in 2005
  dplyr::mutate(year = ifelse(year == 2009, 2005, year)) %>%
  dplyr::filter(year %in% yrs) %>%
  
  ## the template can limit the years included 
  merge(x = temp_eora, 
        y = ., 
        by.x = c('year', 'iso3', 'to'), 
        by.y = c('year', 'iso3', 'to'), all.x = T) %>%
  as.data.frame()



### To generate the same format as Eora in Matrix
df_formatted <- dfs_flow_f %>%
  ungroup() %>%
  arrange(year, iso3, to) %>%
  dplyr::select(year, ctr, everything()) %>%
  spread(key = to, value = value) %>%
  arrange(year, iso3)


### save to the `eora_cleaned` folder --------------------------------------------------------------
df_formatted %>%
  group_by(year) %>% tally()
fname <- paste0(dir.eora_cleaned, ind, '_matrix.xlsx'); fname
writexl::write_xlsx(x = df_formatted, path = fname)
```




```{r - WB data for comparision}
### data was compiled via `10_country_profileData_prepare.Rmd`

### 1. data from WB --------------------------------------------------
xls <- list.files(path = dir.cleaned, pattern = '^Foreign direct investment, net inflows', full.names = T); xls 
inflow_wb_net_in <- readxl::read_excel(path = xls) %>% 
  dplyr::select(-ctr) %>%
  gather(key = 'year', value = 'value', `2000`:ncol(.)) %>%
  dplyr::mutate(value     = as.numeric(value)) %>%  
  dplyr::filter(year == 2015) %>%
  as.data.frame()

### 2. data from IMF -------------------------------------------------
inflow_imf <- df_formatted %>%
  dplyr::filter(year == 2015)
  
source('./Code/func_net_import_direction.R')
inflow_imf_net_in <- func_net_import_direction(df = inflow_imf, direction_i = 1, frac = 0.025) %>%
  dplyr::select(iso3, year, net_in) %>%
  dplyr::rename('net_in_imf' = 'net_in') 


### 3. compare
inflow_net_in_compare <- inflow_imf_net_in %>%
  left_join(x = ., 
            y = inflow_wb_net_in %>%
              dplyr::rename('net_in_wb' = 'value'), 
            by = c('iso3', 'year'))

inflow_net_in_compare %>%
  ggplot() +
  geom_point(aes(x= net_in_imf, y= net_in_wb)) +
  geom_abline(slope=1, color = 'red') +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  theme_bw()
```


