---
title: "FDI"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---


# Set up
```{r Paths and packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
remove(list = ls())

### set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
getwd()

### data path 
source('./Code/_path of data.R')


### packages
source('./Code/_package list.R')
source('./Code/function_scatter_plot.R')

```


# Data

## Ancillary data
```{r}

### extended country-iso3 pair list for matching
load('./Data/Ancillary_Data_iso_eora_ex.RData')  ## ## iso_eora_ex


### use the Eora data as the template 
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')
load(f) # `temp_eora`
```



## Inventor migration 

  Information on inventors’ mobility (Flows), defined as the number of countries a patent inventor changes during lifetime every time she files for a new patent, is obtained from the *World Intellectual Property Organization (WIPO) Database*.
  
  File 1 - "1. Bilateral flows" Bilateral migration corridors, annual data
  
  List of variables in WIPO-BD (release 1.0):
 
  `iso2_ori` - "Country/territory of birth/origin of the inventors" ISO 2-alpha country/area codes. 
  `iso2_des` - "Country/territory of current residence of the migrant inventors"  ISO 2-alpha country/area codes. 
  `prio_year`- "Priority year"	Priority year refers to the first year of the patent application in any patent office
  `flow` - "Flows of inventors" Number of inventors' names, by patent number, with residence (iso2_des) and nationality (iso2_ori), per year
  
  Data range: 1978-2012
  
```{r - preprocessing}
### get the indicator name
ind <- 'Inventor migration'

dir.data <- './Data/data_01_raw/Inventor migration files/'

yrs <- c(2000, 2005, 2010, 2015)

f1 <- paste0(dir.data, 'CSV files/1. Bilateral flows.csv'); f1

d1 <- readr::read_csv(file = f1, show_col_types = F) %>%
  filter(prio_year >= 1990) %>%
  as.data.frame()
names(d1)


df <- d1 %>%
  pivot_wider(names_from = 'prio_year', values_from = 'flow') %>%
  mutate(`2015` = `2012`)

```


```{r - country code}

library(countrycode)

f <- paste0(dir.data, 'CSV files/12. List of countries or territories, and classifications.csv'); f
c <- readr::read_csv(file = f, show_col_types = F) %>%
  select(1:2)
names(c)

## convert country code using https://github.com/vincentarelbundock/countrycode
ctr_code <- d1 %>%
  distinct(iso2_ori, iso2_des) %>%
  gather(key = 'iso', value = 'iso2') %>%
  distinct(iso2) %>%
  filter(!is.na(iso2)) %>%
  ## convert 2c to 3c
  # mutate(iso3c = countrycode(sourcevar = iso2, origin = 'iso2c',  destination = 'iso3c')) %>%
  left_join(x = .,
            y = c,
            by=c('iso2' = 'iso_alpha2_code')) %>%
  rename(iso3c = iso_alpha3_code) %>%
  arrange(!is.na(iso3c))

names(ctr_code)

```


```{r - final 📁}

dfl_ <- df %>%
  left_join(x = .,
            y = ctr_code,
            by=c('iso2_ori' = 'iso2')) %>%
  dplyr::rename("o_iso" = "iso3c") %>%
  select(1:iso2_ori, o_iso, everything()) %>%
  left_join(x = .,
            y = ctr_code,
            by=c('iso2_des' = 'iso2')) %>%
  dplyr::rename("d_iso" = "iso3c") %>%
  select(1:iso2_des, d_iso, everything())



## format as the same as EORA data
df_m <- dfl_ %>%
  dplyr::select(-c(iso2_ori, iso2_des)) %>%
  filter(!is.na(o_iso), !is.na(d_iso)) %>%
  arrange(o_iso, d_iso)

# Target years to fill
target_years <- yrs

df_m_filled <- df_m %>%
  ## to re-sort years in column
  gather(key = 'year', value = 'value', `1990`:ncol(.)) %>%
  dplyr::mutate(year = as.numeric(year)) %>%
  dplyr::filter(year %in% yrs) %>%
  ## the template can limit the years included
  merge(x = temp_eora,
        y = .,
        by.x = c('year', 'iso3', 'to'),
        by.y = c('year', 'o_iso', 'd_iso'), 
        all.x = T) %>%
  as.data.frame()



### To generate the same format as Eora in Matrix
df_m_formatted <- df_m_filled %>%
  ungroup() %>%
  as.data.frame() %>%
  arrange(year, iso3, to) %>%
  dplyr::select(year, ctr, everything()) %>%
  pivot_wider(names_from = to, values_from = value) %>%
  arrange(year, iso3)


# ## check duplicates 
# df_m_filled %>%
#   dplyr::group_by(year, ctr, iso3, to) %>%
#   dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
#   dplyr::filter(n > 1L)

df_m_formatted %>%
  group_by(year) %>% tally()



### save to the `eora_cleaned` folder ----------------------------------------------------
fname <- paste0(dir.eora_cleaned, ind, '_matrix.xlsx'); fname
writexl::write_xlsx(x = df_m_formatted, path = fname)
```





