---
title: "FDI"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---


# Set up
```{r Paths and packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### To clear your environment 
remove(list = ls())

### set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
getwd()

### data path 
source('./Code/_path of data.R')


### packages
source('./Code/_package list.R')
source('./Code/function_scatter_plot.R')

```


# Data

## Ancillary data
```{r}

### extended country-iso3 pair list for matching
load('./Data/Ancillary_Data_iso_eora_ex.RData')  ## ## iso_eora_ex


### use the Eora data as the template 
f <- paste0('./Data/data_02_intermediate/dt02_flows/', 'template_eora.RData')
load(f) # `temp_eora`
```



## CPIS

  The *Coordinated Portfolio Investment Survey* (CPIS) is a voluntary data collection exercise conducted under the auspices of the IMF. 
  The CPIS is a voluntary data collection exercise that assembles data on countries' international holdings of equities and long- and short-term debt securities.39 Countries report data on (i) their holdings of portfolio investment assets issued by residents in other countries and (ii) their portfolio investment liabilities issued by domestic residents and held by residents in other countries. The data are reported at the bilateral country-to-country level for all sectors. 

```{r - preprocessing}
### get the indicator name
ind <- 'CPIS'

library(countrycode)

yrs <- c(2000, 2005, 2010, 2015)

cc <- load(paste0(dir.cpis,'codelist_panel.rda'))

f <- paste0(dir.cpis, 'CPIS_12-16-2024 19-54-07-13_timeSeries.csv'); f

df <- readr::read_csv(file = f, show_col_types = F) %>%
  dplyr::rename(
    o_name = `Country Name`, 
    o_code = `Country Code`,
    d_name = `Counterpart Country Name`, 
    d_code = `Counterpart Country Code`
  ) %>%
  select(-contains('Sector'))  %>%
  select(1:`2023`) %>%
  filter(!`Indicator Code` %in% c('I_L_T_T_T_BP6_DV_USD', 'I_L_T_T_T_BP6_USD'))

names(df)

# unique(df$`Counterpart Sector Name`)
unique(df$Attribute)
distinct(df, `Indicator Name`, `Indicator Code`)

df_ <- df %>%
  filter(Attribute == "Value") %>%
  select(-Attribute) %>%
  filter(`Indicator Code` %in% c('I_A_T_T_T_BP6_USD'))

distinct(df_, `Indicator Name`, `Indicator Code`)

dfl <- df_ %>%
  as.data.frame() %>%
  select(-contains('Indicator'))  %>%
  mutate_if(is.numeric, as.character) %>%
  pivot_longer(cols = `1997`:`2023`, names_to = 'year', values_to = 'value') %>%
  mutate(value = as.numeric(value)) %>%
  ## check data 
  # filter(o_name == 'Germany', year == 2020) %>%
  as.data.frame()
```


```{r - preprocessing - country code}

ctr_list1 <- df %>% distinct(o_name, o_code) %>%
  dplyr::rename(name = o_name, 
                code = o_code)
ctr_list2 <- df %>% distinct(d_name, d_code) %>%
  dplyr::rename(name = d_name, 
                code = d_code)

## convert country code using https://github.com/vincentarelbundock/countrycode
ctr_list <- rbind(ctr_list1, ctr_list2) %>%
  distinct() %>%
  mutate(
    code = as.numeric(code),
    iso = countrycode(code, origin = 'imf', destination = "iso3c")) %>%
  select(1:code, iso, everything()) %>%
  arrange(!is.na(iso))


## clean the data 
iso3_code <- iso_eora_ex %>% dplyr::select(-IMF)
  

ctr_list_na <- ctr_list %>%
  filter(is.na(iso)) %>%
  select(-iso) %>%
  dplyr::mutate(name = gsub('\\.|\\:|\\,', '', name)) %>%
  left_join(x = .,
          y = iso3_code,
          by=c('name' = 'Row')) %>%
  dplyr::rename("iso" = "iso3_eora") %>%
  arrange(!is.na(iso)) 

ctr_code1 <- ctr_list %>% filter(!is.na(iso))
ctr_code2 <- ctr_list_na %>% filter(!is.na(iso))
ctr_code <- rbind(ctr_code1, ctr_code2) %>%
  select(-name)
```


```{r - final 📁}

dfl_ <- dfl %>%
  mutate(o_code = as.numeric(o_code),
         d_code = as.numeric(d_code) ) %>%
  left_join(x = .,
            y = ctr_code,
            by=c('o_code' = 'code')) %>%
  dplyr::rename("o_iso" = "iso") %>%
  select(1:o_code, o_iso, everything()) %>%
  left_join(x = .,
            y = ctr_code,
            by=c('d_code' = 'code')) %>%
  dplyr::rename("d_iso" = "iso") %>%
  select(1:d_code, d_iso, everything())



## format as the same as EORA data
df_m <- dfl_ %>%
  ## put year as column and fill na if needed
  pivot_wider(names_from = 'year', values_from = 'value') %>%
  ## fill data gap using adjacent years
  dplyr::mutate(
    `2000` = `2001`
  ) %>%
  dplyr::select(1:`1997`, `2000`, everything())



# Target years to fill
target_years <- yrs

# Function to fill NA values in target year columns
fill_na_adjacent_cols <- function(df, target_years) {
  for (year in target_years) {
    year <- as.character(year)  # Ensure year is a string
    
    # Find the previous and next year columns
    year_next <- as.character(as.numeric(year) + 1)
    year_prev <- as.character(as.numeric(year) - 1)
    
    # Check if these columns exist in the data
    if (year_next %in% colnames(df)) {
      df[[year]][is.na(df[[year]])] <- df[[year_next]][is.na(df[[year]])]
    }
    if (year_prev %in% colnames(df)) {
      df[[year]][is.na(df[[year]])] <- df[[year_prev]][is.na(df[[year]])]
    }
  }
  return(df)
}

# Apply the function to the data frame
df_m_filled <- fill_na_adjacent_cols(df_m, target_years) %>%
  dplyr::select(-c(o_name, d_name, o_code, d_code)) %>%
  pivot_longer(cols = `1997`:`2023`, names_to = 'year', values_to = 'value') %>%
  dplyr::mutate(year = as.numeric(year)) %>%
  dplyr::filter(year %in% yrs) %>%
  ## the template can limit the years included
  merge(x = temp_eora,
        y = .,
        by.x = c('year', 'iso3', 'to'),
        by.y = c('year', 'o_iso', 'd_iso'), all.x = T) %>%
  as.data.frame()



### To generate the same format as Eora in Matrix
df_m_formatted <- df_m_filled %>%
  ungroup() %>%
  arrange(year, iso3, to) %>%
  dplyr::select(year, ctr, everything()) %>%
  spread(key = to, value = value) %>%
  arrange(year, iso3)

df_m_formatted %>%
  group_by(year) %>% tally()

### save to the `eora_cleaned` folder ----------------------------------------------------


fname <- paste0(dir.eora_cleaned, ind, '_matrix.xlsx'); fname
writexl::write_xlsx(x = df_m_formatted, path = fname)
```





